---
title: "Predator phylogenetic diversity decreases predation rate via antagonistic interactions"
author: "A. Andrew M. MacDonald; Diane S. Srivastava; Gustavo Q. Romero"
fontsize: 12pt
output:
  pdf_document:
    includes:
      in_header: myheader.tex
    latex_engine: xelatex
  md_document: default
  html_document: default
  word_document:
    reference_docx: reference.docx
geometry: margin=1in
bibliography: /home/andrew/Documents/reference/Mendeley/Mendeley_reference_lists/@MS_pdpaper.bib
---

```{r options_packages_data, message=FALSE, echo=FALSE}
## set chunk options
library(knitr)
opts_chunk$set(message=FALSE,warning=FALSE,echo=FALSE, fig.cap=TRUE,
               cache = TRUE, cache.path = "./cache/",
               dev='pdf', fig.path = './figure/')

## load packages
library(ggplot2)
library(bipartite)
library(vegan)
library(picante)
library(pander)
library(dplyr)
library(magrittr)
library(xtable)
library(stargazer)
library(tidyr)
library(reshape2)

## ggplot2 theme
mytheme <- theme_bw()

# occurrance data -- abundance
occur <- read.csv("../data/reorganized_data/predator.cooccur.txt",
                  stringsAsFactor=FALSE) %>%  tbl_df()
# occurrance data -- metabolic capacity
metabolic  <- read.csv("../data/reorganized_data/predator.cooccur.metabolic.txt",
                       stringsAsFactor=FALSE) %>%  tbl_df()
# feeding trial data
feeding_trials <- read.csv("../data/reorganized_data/reorganized.feeding.trial.data.csv",
                    stringsAsFactors=FALSE) %>%   tbl_df()
# experimental data
pd <- read.csv("../data/reorganized_data/pd_exp_cleaned_data.csv",
               stringsAsFactors=FALSE) %>%   tbl_df()
# enriched leaves -- ie just the N data for the detritus we put in.
enriched <- read.csv("../data/reorganized_data/enriched_leaves.csv",
                     stringsAsFactors=FALSE) %>%   tbl_df()
# phylogeny data
predtree_timetree_ages <- read.tree("../data/reorganized_data/predator_tree_time.newick")
# randomization tests
rand.means <- read.csv("../data/predator.div.experiment/randomizations.group.means.csv") %>%
  tbl_df()

## load in functions
source("../R.scripts/FUNCTIONS_predator_diversity.R")
```

```{r experiment_data_postprocess}
#dropping a record that seems to have been 90% decomposed!
pd <- pd %>%
  transform(decomp=ifelse(decomp>0.7,NA,decomp))
```

## Introduction

Predators can have strong top-down effects, both on community structure and
ecosystem processes [@Estes2011].  The effects of all predators in a community
(the predator assemblage) is often dependent on species composition, and can be
more or less than the sum of effect sizes of individual predators
[@Sih1998a;@Ives2005]. Detailed studies manipulating predator diversity have
identified many direct and indirect mechanisms of predator interaction. For
example, predators may differ in their preference for microhabitats resulting in
spatial or temporal partitioning between predator species [@Schmitz2007].  Even
when predators occupy the same patch, they may feed on different prey species
and so have additive top-down effects [@Schmitz2007]. Non-additive effects of
predator diversity occur when predators interact directly with each other, or
indirectly via prey species, in a way that increases or decreases total
predation. For example, antagonistic predator-predator interactions can reduce
the top-down effects of a diverse consumer assemblage, e.g. when
predators feed directly on each other (intra-guild predation) or modify the
behaviour of predators or prey via non- consumptive (i.e. trait-mediated)
interactions [@Sih1998a;@Griswold2006;@Nystrom2001].  While there are many
mechanisms potentially underlying predator compositional effects, we lack a
means of predicting _a priori_ which combinations are likely to have
antagonistic versus synergistic effects on ecosystem functions.

The measurement of phylogenetic diversity has become a popular means for
ecologists to make inferences about ecological and evolutionary mechanisms
behind observed patterns in natural communities [@Cavender-Bares2009]. For example, phylogenetic
diversity of plant communities is better correlated with productivity than is
species richness or diversity [e.g. @Cadotte2009;@Cadotte2008;@Godoy2014].
Central to most ecological applications of phylogenetic diversity is the
assumption that increased phylogenetic distance implies increased ecological
dissimilarity -- either in the form of differences in species niches,
interactions, or functional traits. When these relationships occur, the
resulting variation in ecologically-relevant traits and interactions could
increase ecosystem function [@Srivastava2012c]. Despite the prevalence of
phylogenetic community ecology, and the importance of predators to natural
systems, the phylogenetic diversity of local predator assemblages has rarely
been measured [@Bersier2008;@Naisbit2011]. Many studies of phylogenetic signal
in predator traits focus on whole clades, rather than local assemblages (e.g.
*Anolis* lizards [@Knouft2006], warblers [@Bohning-Gaese2003], treeboas
[@Henderson2013] and wasps [@Udriene2005]) making it difficult to connect these
results to top- down effects at the scale of a local community. While these
clade-specific studies often find weak evidence for phylogenetic signal in
ecologically- relevant traits, studies at the level of the whole biosphere
[@Gomez2010;@Bersier2008] demonstrate that related organisms often have similar
interspecific interactions -- i.e. related predators often consume similar prey.
At the level of a community, the effect of a predator assemblage will depend on
both the spatial distribution and trophic interactions of all predator species.
Phylogeny may constrain species distributions when distant relatives have
distinct fundamental niches, while close relatives are too similar to co-exist
[@Webb2002;@Emerson2008]. The effects of other predators can also be
be important: during dispersal, predators may use
cues of prey, competitors, or intra-guild predators when selecting a habitat patch
[@Chase2009a;@McCauley2008a;@Srivastava2008]. After colonization, patterns of
co-occurrence may be further modified by intraguild predation
[@Lounibos2008;@Juliano2009] or competition [@Fincke1997].

When predators do co-occur, their direct effects on each other and on their prey is determined by their diet breadth. Diet breadth (the diversity of prey species which predators consume) will depend on their feeding traits and nutritional -- all of which may be phylogenetically conserved. If this is the case, then predator assemblages with higher phylogenetic diversity will show greater complementarity and a stronger top-down effect on ecosystem function
[@Finke2008a]. For
example, some predators are gape-limited, and can swallow any prey smaller than
their mouth; these predators typically broaden their diet as they grow (e.g.,
larvae of Odonata) [@Werner1984].  Such ontogenetic changes in diet breadth can
radically influence food web structure [@Moya-Larano2011]. Other
predators (e.g. Leeches, or Diptera:Tabanidae) are "piercing-sucking" predators
[@Bay1974] and may be more general consumers at all life stages, since they can
feed on larger prey individuals even at very early predator life stages.  In
some cases, predator diets may extend to include other predators, leading to
direct negative interactions such as intraguild predation which may also have a
phylogenetic signal [@Pfennig2000].

Indirect interactions among predators are common in many communities, and can lead to
nonlinearities which prevent the inference of predator effect from the additive
combination of individual predators [@Sih1998a]. For example, decreasing
predator richness has been shown to increase herbivory in a three-level kelp
food web [@Byrnes2006]. The complex traits that underlie such species
interactions ("higher-order" interactions, Sih *et al.*
[-@Sih1998a]) may also correlate with phylogeny; if so, then phylogenetic diversity will be a better predictor of ecosystem function than richness alone [@Cadotte2009]. For example,
phylogenetically diverse assemblages of mycorrhizae increased primary
productivity more than equally speciose treatments from the same lineage
[@Maherali2007]. The effect of predators on prey may also be non-additive, the
result of indirect interactions either with prey species or other predators:
prey species may respond to the presence of one predator by a behavioural shift
that increases the predation by another predator population [@Carey2010a]. These
trait-mediated indirect effects (TMII, [@Werner2003]) can be difficult to predict;
however phylogenetic relatedness has been suggested as a possible means of
predicting the combined effect of predators [@Naisbit2011]

We used a series of observations, lab feeding trials, and manipulative field
experiments to measure how the phylogenetic diversity of the predator assemblage
predicts community composition and ecosystem function.  We test three related
hypotheses concerning co-occurance patterns, diet similarity and top-down
ecosystem effects of diverse predators, using a natural mesocosm: the community
of invertebrates living within bromeliads. Bromeliads (Bromeliaceae) are
epiphytic plants native to the Neotropics. Many bromeliad species contain water, detritus
and a complex insect food web within their leaves; the decomposition of this
detritus supplies nutrients for the bromeliad [@Benzing2000]. The small size of
these habitats permits direct manipulations of entire food webs, manipulations
which would be difficult in most natural systems.  Within this aquatic food web, damselfly larvae (e.g. *Leptagrion* spp., Odonata:Coenagrionidae) are important predators;
their presence dramatically alters community dynamics (e.g. decreases
insect emergence [@Starzomski2010] and increases nutrient cycling
[@Ngai2006]). 

1. *species co-occurrence*: closely-related predators may occur together more frequently than less-related predators if there is a strong phylogenetic signal to habitat requirements.   Alternatively, very closely related species may never co-occur because high overlap in ecological niches results in competitive exclusion.

2. *diet similarity*: similarity in diet (as measured by feeding trials)
decreases with phylogenetic distance if diet is phylogenetically conserved. Alternatively,  closely related species may have evolved different diets to allow coexistence.

3. *ecosystem-level effects*: Our experiments at the level of the whole habitat patch (i.e. a single bromeliad) allows us to examine direct and indirect effects of predator combinations.
    a) monoculture treatments allow us to assess the effect of each predator on a complete prey community and ecosystem function.  Phylogenetic similarity among predators is predicted to  correlate with similarity in single-species effects if related predators share similar trophic interactions (e.g. predation rate, diet breadth)
    b) By comparing treatments with pairs of predators to treatments which received a monoculture of each predator, we are able to estimate additive and nonadditive effects.  If distance between species correlates with increasing difference, we will see an increase in these non-additive effects with phylogenetic distance. At the extreme, differences between predators may lead to IGP among predators.

## Methods

## Site and Species

We conducted all observations and experiments in Parque Estadual da Ilha do
Cardoso ($25^{\circ} 03^{\prime}$ S, $47^{\circ}53^{\prime}$ W), a 22.5 ha
island off the south coast of São Paulo state, Brazil. We worked in a closed
coastal forest (*restinga*) the understory of which is mostly covered by
*Quesnelia arvensis* Mez. (Bromeliaceae), a large terrestrial bromeliad that
accumulates up to 2.8 L of rainwater in tanks formed by individual leaves. We used two datasets: an observational survey of insect diversity from 2008, and a manipulative experiment in 2010. Our observational survey found more than 47 species of macroinvertebrates in these aquatic communities [@Romero2010].
This diversity encompasses multiple trophic and functional groups: filter feeders (Diptera:Culicidae); detritivores including shredders
(Diptera:Tipulidae, Trichoptera), scrapers (Coleoptera:Scirtidae), and collectors
(All Diptera:Chironomidae, Syrphidae, Psychodidae). These species are the main prey for a 
diverse predator assemblage of at least three species of damselfly larvae (_Leptagrion_ spp., Zygoptera:Coenagrionidae), two species of predatory Horse Fly larvae (Diptera:Tabanidae), and 2
species of leech (Hirudinidae).  Other, smaller-bodied predators present include predatory midge larvae (Diptera:Chironomidae:Tanypodinae) and less abundant species such as Dytiscid beetles.

### Phylogenetic distance, metabolic capacity and niche overlap

```{r CALC_nodeages}

nodeages <- list.files(path="../data/TreeData",pattern="*.csv",full.names=TRUE) %>%
  data.frame(paths = .,stringsAsFactors = FALSE) %>%
  mutate(nodename = basename(paths) %>% gsub(".csv","",.)) %>%
  group_by(nodename) %>%
  do(age_data = read.csv(.$paths)) %>%
  mutate(Nstudies = nrow(age_data),
         minT = min(age_data %>% extract2("Time")),
         maxT = max(age_data %>% extract2("Time")))


## move to supp mat
phylogeny_node_age <- nodeages %>%
  select(nodename,age_data,Nstudies) %>% 
  do(extract2(.,"age_data") %>% as.data.frame(stringsAsFactors = FALSE)) %>%
  lapply(.,function(x) gsub(pattern = "\\(.+\\)",replacement = "",x)) %>% 
  data.frame

# function to print out range

agerange <- function(Name,pna = nodeages){
  minT <- pna %>%
    filter(nodename == Name) %>%
    extract2("minT")
  
  maxT <- pna %>%
    filter(nodename == Name) %>%
    extract2("maxT")
  
  paste0(minT," to ",maxT)
  }

nstudy <- function(Name, dat = nodeages){
  dat %>%
    filter(nodename == Name) %>%
    extract2("Nstudies")
  }

              
n.nodes <- nrow(nodeages)
nstudies <- sum(nodeages$nstudies)

# nodeages <- lapply(list.files(path="../data/TreeData/",pattern="*.csv",full.names=TRUE),read.csv)
# names(nodeages) <- list.files(path="../data/TreeData/",pattern="*.csv")
# nstudies <- sapply(nodeages,nrow)
# #names(nodeages)[which(nstudies>1)]
# n.nodes <- length(nodeages)
```
In 2008, we counted and measured macroinvertebrates in an observational study of `r ncol(metabolic)-1` bromeliads. We
identified `r metabolic$Taxa %>% unique %>% length` species as predators, including two groups of congeneric taxa -- _Bezzia_ sp.
(Diptera:Ceratopogonidae) with two species and _Leptagrion_ sp.
(Odonata:Coenagrionidae) with three -- and two of confamilial taxa -- three species of
Tabanidae and two of Empididae, all Diptera.  Deeper divisions were also present: three families of Diptera are represented by a single predator species each (Dolichopodidae, Corethrellidae and
Chironomidae) and the deepest taxonomic divide is between all insects present and a
species of predatory leech (Annelida:Hirudinidae).

We obtained node age estimates for all seven internal nodes of the tree, using `www.timetree.org`, an online database of published  molecular time estimates [@Hedges2006].  Most dates came from only a single study; where multiple dates were found we used the median estimate: Insecta--Hirudina (`r agerange("insects.to.leeches")` Mya, n=`r nstudy("insects.to.leeches")` studies), Odonata--Tabanidae (`r agerange("odonata-Tabanidae")` Mya, n=`r nstudy("odonata-Tabanidae.csv")` studies) and Tabanidae--Diptera (`r agerange("tabanidae_culidicae_ie_Diptera")` Mya, n=`r nstudy("tabanidae_culidicae_ie_Diptera")` studies).  Node age data was available
for all but the youngest nodes, where either a lack of taxonomic
information (e.g. Tabanidae) or a lack of phylogenetic study (e.g. _Leptagrion_)
prevented more information from being included.   These branches were left as
polytomies, and were all assigned identical, arbitrary and short branch lengths
(15 Mya).

Species co-occurrence is often measured in terms of non-random patterns of
species presence/absence or abundance, but such measures will only be poorly related
to the functional effects of species when species differ substantially in body
size.  Integrating the allometric relationship between body size and feeding
rate [@Brown2004; @Wilby2005] over all individuals of a species allows estimates
of "metabolic capacity", or the potential energy requirements of a species
[@Srivastava2009a]. Metabolic capacity is equal to individual body
mass raised to the power of 0.69 (an insect-specific exponent determined by Peters 19xx, [@Chown2007]); this reflects the nonlinearity of feeding rate
on body size across many invertebrate taxa. We used metabolic capacity to inform
both our observational results and our experimental design (details below), with the exception
of our feeding trial data.  This is because the feeding trials were intended to
measure which prey our predators ate, rather than their feeding rate (only the
latter should scale with metabolic capacity).

Questions 1 and 2 above focus on similarity ("niche overlap") between predator species. Question 1 hypothesizes that if habitat preference is phylogenetically conserved and competition between predators unimportant, then distributional similarity will decline with increased phylogenetic distance between a pair of predators. Question 2 hypothesizes that if diet is conserved, then diet similarity declines with increased phylogenetic distance between a pair of predators. We evaluated both distributional and diet similarity between predators using Pianka's index of niche overlap [@Pianka1974]:

$O_{kl}=\dfrac{\sum_i^n{p_{il} p_{ik}}}{\sqrt{\sum_i^n{p_{il}^2} \sum_i^n{p_{ik}^2}}}$

For each pair of predators, $p_{ik}$ and $p_{il}$ represent the preference of
predator $k$ (or $l$) for resource or habitat $i$. In terms of predator co-occurrence, $n$=`r ncol(metabolic)-1` bromeliads surveyed in the observational dataset described below. In terms of diet similarity, the number of resources ($n$) is defined as the total number of prey
species assayed with both predator taxa, and preference is defined as the
proportion of diet trials (see below) for each predator-prey combination that resulted in prey
mortality.  

We quantified the effect of phylogenetic distance on each of distributional and diet similarity.  First, we calculated phylogenetic distance between each pair of species, then fit several functions to this relationship (linear, constant, and several appropriate nonlinear functions).  We compared these models using AIC, selected the best model, and generated confidence intervals as appropriate (parametric or bootstrap for linear and nonlinear, respectively).

###  Distributional Similarity 

We used the 2008 survey to examine patterns of co-occurrence among predator taxa.  Each bromeliad was dissected and washed to remove invertebrates.  The
resulting water was filtered through two sieves (250 and 850 µm), which removed
particulate organic matter without losing any invertebrates.  All invertebrates
were counted and identified to the lowest taxonomic level possible. The body
length of all individuals was measured, when possible for small and medium-sized taxa (< 1cm final instar) and always for large-bodied taxa (> 1 cm final instar).

We calculated total metabolic capacity of each predator species in each
bromeliad by summing estimates of biomass for all individuals of the same
species within a plant.  We used these values to calculate niche overlap as
measured by habitat occupancy.  Overlap was calculated for each pair of
predators across all bromeliads, even if neither species occurred there.

### Diet Similarity
```{r CALC_feeding_trials}
##  predator x prey trials


## numbers for text
ntrials <- sum(feeding_trials$number.trials)
npred <- length(unique(feeding_trials$predator.names))
nprey <- length(unique(feeding_trials$Prey.species))
ncombos <- nrow(feeding_trials)
rep_range <- range(feeding_trials$number.trials)
```

We conducted `r ntrials` feeding trials of `r npred` predator taxa fed `r nprey` prey taxa between March and April 2011.  We  covered all potential predator-prey pairs present in the experiment
(described below), and attempted to perform all other trials whenever possible. However, due to the rarity of some taxa many predator-prey pairs were not possible; we tested `r ncombos` pairwise combinations.  Most trials were replicated at least five times, but the number of replicates for various combinations ranged from `r rep_range[[1]]` to `r rep_range[[2]]`. We placed
predators together with prey in a 50ml vial, with a leaf or stick for
substrate. The only exception was the tabanid larvae, which we
placed between two vertical surfaces to imitate the narrow space found in
bromeliad leaf axils, their preferred microhabitat.  Generally our trials contained a single predator and a
single prey individual, except in the case of very small prey (_Elpidium_ sp.)
or predators (_Monopelopia_ sp.) in which case we increased the density.  We
replicated each combination up to five times where possible, and allowed one day for
predation to occur.

Again, we estimated overlap between different predators using Pianka's index,
using the proportion of trials which resulted in predation as our measure of
predator preference. Because not every predator/prey pair was equally
replicated, we weighted our analysis by  the number of trials conducted.

### Community effect experiment

Our experiment was created with two goals in mind: first, to measure the major
effects of each of these predators on their prey and ecosystem function and second, to estimate the non-
additive effects of pairwise predator combinations.  The strength of these non-additive effects could then be related to the phylogenetic distance between a pair of predator species.

We selected the four most common predators in this system, which also created a
range of relatedness: two congeneric damselflies (*Leptagrion andromache* and
*Leptagrion elongatum*), two insects (*L. elongatum* and a Tabanid predatory
fly), and two invertebrates (*L. elongatum* and leeches).  We used each of the
pairs of predators just described, as well as each species in monoculture, and a
predator-free control (8 treatments, n=5).  Combinations were substitutive,
maintaining the same amount of total predator metabolic capacity to isolate the
effects of predator behaviour and traits. Substitutive experiments often hold
total abundance constant, but when species differ substantially in body size -
as in this experiment - allometric effects of body size on feeding rate can
confound detection of effects based on trophic traits or species
interactions, and standardizing to community metabolic capacity is preferred
[@Srivastava2009a]. This experiment allows the
estimation of the effect of each predator species (monoculture treatments), as
well as the detection of non- additive effects in predator combinations.

In Feburary 2011, bromeliads between 90 and 200ml in capacity were collected, thoroughly
washed to remove organisms and detritus and soaked for 12 hours in a tub of water.  They were then hung for 48
hours to dry.  One bromeliad dissected after this procedure contained no
insects.

Each bromeliad was supplied with dried leaves, simulating natural detritus
inputs from the canopy.  We enriched these leaves with ^15^N by fertilizing five
(Jabuticaba, *Plinia cauliflora*) plants with 40ml pot^-1^ day^-1^ of 5g L^-1^
ammonium sulphate containing 10% atom excess of ^15^N over 21 days. Whole leaves
were then picked from plants and air-dried until constant weight, and then
soaked for three days and the water discarded. About 1.5 g of leaves were placed
in each bromeliad (`r mean(pd$mass.g.)`g ± `r sd(pd$mass.g.)`).

Each bromeliad was stocked with a representative insect community. The densities
of each prey taxon were calculated from the 2008 observational dataset, using
data from bromeliads of similar size to those in our experiment.  All densities
used were within the range of these calculated abundances, and all experimental
bromeliads received the same insect community.  Halfway through the experiment,
insects were added to bromeliads a second time to simulate the continuous
oviposition that characterizes the system.  Throughout the experiment, all bromeliads were enclosed
with a mesh cage topped with a malaise trap and checked daily for emergence of
adults.

In this experiment we measured five response variables: production of fine
particulate organic matter (FPOM), decomposition of coarse detritus, bromeliad
growth, uptakle of detrital nitrogen into bromeliad tissue, and survival of invertebrate prey (emerged adults +
surviving larvae). We analyzed each of these responses with ANOVA.  For each pair of predator species and each response type, we
calculated the non-additive effect as the difference between the
response in bromeliads with both predator species (n=5) and the mean response in bromeliads with either one of these two predator species 
(n=5 for each predator species).  We generated bootstrap confidence intervals for
these nonadditive effects; confidence intervals which do not overlap zero
indicate a significant non-additive effect of a predator combination.



## Results

### Distributional similarity

```{r CALC_phylogenetic_distance}
####### metabolic matrix ####
## we need to calculate two distance matrices:
## 1) metabolic capacity distance
## 2) phylogenetic distance

## metabolic matrix -- the "distance" between predator co-occurance, measured as metabolism

####### phylogeny matrix ####
## Calculate distances
phylogenetic_distance  <- predtree_timetree_ages %>%
  cophenetic() %>%
  matrix_to_df() %>%
  select(phylopred1=Var1,
         phylopred2=Var2,
         phylodistance=value) %>%
  mutate(pairs_RH=paste0(phylopred1,"_",phylopred2),
         pairs_LH=paste0(phylopred2,"_",phylopred1)
         ) %>%
  melt(id.vars=c("phylopred1","phylopred2","phylodistance"),
       value.name="species_pair",
       variable.name="L_or_R") 
## my 

```

```{r CALC_metabolic_distance}

metabolic_distance <- paired_predator_pianka(pred_x_resource = metabolic,
                                             pred_colname = "Taxa",... =-Taxa)

```

```{r CALC_diet_overlap}
## When calculating similiarity, we start with a predator (row) by resource (column) dataframe.
# pred x prey matrix, with cells = proportion eate

# feeding_trials %>% arrange(predator.names) %>% mutate(prop.eaten=eaten/number.trials)

prop.eaten <- feeding_trials %>% 
  filter(!Prey.species %in% c("Leptagrion andromache <5mm",
                              "Leptagrion.elongatum")) %>% 
  filter(!predator.names %in% c("Phylloicus bromeliarum", "Dysticid")) %>%
  mutate(prop.eaten=eaten/number.trials) %>%
  arrange(predator.names,Prey.species) %>%
  select(-number.trials,-eaten) %>%
  spread(Prey.species,prop.eaten,fill = NA)
   #dcast(predator.names~Prey.species,value.var="prop.eaten",fun.aggregate = mean)

write.csv(prop.eaten,"./analyses/proportion.eaten.csv")

diet_overlap <- paired_predator_pianka(pred_x_resource = prop.eaten,
                                       pred_colname = "predator.names",
                                       -predator.names)

```

```{r CALC_summarize_random_test}
## summarize randomization test results
betternames <- c("elong + andro" = "Leptagrion.elongatum_Leptagrion.andromache",
                 "elong + tab" = "Leptagrion.elongatum_Tabanidae.spA",
                 "elong + leech" = "Leptagrion.elongatum_Hirudinidae")
                 
summarize_random_test <- rand.means %>% 
  # remove annoying X column
  select(-X) %>%
  # melt, so that all responses can be summarized at the same time
  gather(response_var,value,-sp.pair) %>% 
  group_by(sp.pair,response_var) %>% 
  summarise(mean=mean(value),
            lower=quantile(value,probs=c(0.025)),
            upper=quantile(value,probs=c(0.975))
            ) %>%
  # sequence of increasing PD
  ungroup() %>%
  mutate(species_pair = betternames[sp.pair] %>% as.character)
```

```{r DATA_merge}
## we need to merge together several matrices:
## metabolic occurance + predator phylogenetic distance
## diet similarity + predator phylogenetic distance
## experiment randomization results + predator phylogenetic distance

## note that the nomeclature of the columns keeps `sp.pair` as the only shared name among columns.
metabolic_occur_phylo <- metabolic_distance %>%
  left_join(phylogenetic_distance)

diet_overlap_phylo <- left_join(diet_overlap,phylogenetic_distance)

summarize_randoms_phylo <- left_join(summarize_random_test,phylogenetic_distance)

```

```{r CALC_metabolic_numbers}
## numbers for text:
meta_range <- metabolic %>%
  melt(id.vars="Taxa") %>%
  select(-variable) %>%
  filter(value>0) %>%
  filter(value%in%range(value)) %>%
  transform(value=signif(value,2)) %>%
  unique()

smallsp <- meta_range[1,]
largesp <- meta_range[2,]

nspp <- metabolic %>%
  melt(id.vars="Taxa",variable.name="brom") %>%
  filter(value>0) %>% 
  group_by(brom) %>%
  summarise(nspp=n()) %>%
  summarise(meanpred=mean(nspp),
            sdpred=signif(sd(nspp),2)) 

```

```{r TEST_metabolic~phylo}
meta_phylo_lm_summary <- with(metabolic_occur_phylo,lm(overlap~phylodistance)) %>%
  summary()

## caculate summaries 
pval_metabol <- round(pf(meta_phylo_lm_summary$fstatistic[1],
                         meta_phylo_lm_summary$fstatistic[2],
                         meta_phylo_lm_summary$fstatistic[3],lower.tail=FALSE)
                      ,digits=2)
```

Across all bromeliads, predator species differed widely in metabolic capacity, from `r smallsp$value` for a species of _Monopelopia_ to `r largesp$value` for large predatory flies in the family Tabanidae.  Predators often co-occurred in bromeliads (`r nspp$meanpred` ± `r nspp$sdpred` predator species per plant).  However, the niche overlap between the total metabolic capacity of pairs of predators did not show any relationship with phylogenetic distance between them (F~`r meta_phylo_lm_summary$fstatistic[["numdf"]]`,`r meta_phylo_lm_summary$fstatistic[["dendf"]]`~=`r meta_phylo_lm_summary$fstatistic[["value"]]`, p=`r pval_metabol`).

### diet similarity

```{r CALC_diet_numbers}
## Diet numbers for text:
## what percentage of total trials resulted in predation?
percentpredation <- feeding_trials %>%
  transform(percent.eaten=eaten/number.trials) %>%
  group_by(predator.names) %>%
  summarise(mean.pred=mean(percent.eaten),
            totaltrials=sum(number.trials),
            nprey=n())
  
andro <- percentpredation %>%
  filter(predator.names=="Leptagrion.andromache") %>%
  select(mean.pred) %>%
  transform(mean.pred=round(mean.pred*100,1)) %>%
  as.numeric()

elong <- percentpredation %>%
  filter(predator.names=="Leptagrion.elongatum") %>%
  select(mean.pred) %>%
  transform(mean.pred=round(mean.pred*100,1)) %>%
  as.numeric()
  
```

```{r TEST_diet-sim_as_PD,results='asis'}
## test a squared term with 
# 
# diet_overlap_phylo %>%
#   #filter(nspp>1) %>%
#   with(lm(overlap~phylodistance,weights=nspp)) %>%
#   pander(caption="Linear model of diet overlap as a function of phylogenetic distance between predators.")

#summary(dietoverlap_lm)

diet_phylo_lm_summary <- diet_overlap_phylo %>%
  #filter(nspp>1) %>%
  with(lm(overlap~phylodistance,weights=nspp)) %>%
  summary()



# diet_phylo_lm_summary <- diet_overlap_phylo %>%
#   #filter(nspp>1) %>%
#   glm(overlap~phylodistance,weights=nspp,data=.,family=binomial) %>%
#   summary()


dietmat <- diet_overlap_phylo %>%
  select(-species_pair) %>%
  unique() %>%
  dcast(phylopred1~phylopred2,value.var="overlap") %>%
  data.frame(row.names="phylopred1") %>%
  as.matrix()

```


Among the most common predator taxa (i.e. those used in our experiment, described below) the damselflies (_Leptagrion andromache_ and _Leptagrion elongatum_) showed the highest rates of prey consumption (prey consumed in `r andro`% and `r elong`% of trials, respectively). All predators showed a very generalist diet breadth, consuming nearly all species offered to them.  However, more phylogenetically distant predators differed in their preference of prey species, as measured by the niche overlap index (F~`r diet_phylo_lm_summary$fstatistic[["numdf"]]`,
`r diet_phylo_lm_summary$fstatistic[["dendf"]]`~=`r round(diet_phylo_lm_summary $fstatistic[["value"]],2)`, p=`r round(pf(diet_phylo_lm_summary $fstatistic[1],diet_phylo_lm_summary $fstatistic[2],diet_phylo_lm_summary $fstatistic[3],lower.tail=FALSE),digits=3)`, regression weighted by the number of prey species assayed.)

### Ecosystem-level effects and phylogenetic distance

```{r experimentsummary}
predeffect <- function(resp="total.surv"){
  diffeffect <- (mean(pd[[resp]][pd$treatment!="control"],na.rm=TRUE)-mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE))/mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE)
  round(diffeffect,digits=2)*100
  }

# 
# ddply(pd,.(treatment),summarize,meansurv=mean(total.surv))
# 
# mean(pd$emerged)
# mean(rowSums(pd[c("Culicidae","Chironomidae","Tipulidae","Scirtidae")])
```

In our manipulative experiment, we placed a standardized prey community into bromeliads and measured five response variables. Predators had a large effect on prey survivorship: on average all predator treatments showed `r abs(predeffect())`% lower prey emerging or surviving as larvae relative to the predator-free control.  Nitrogen transport to bromeliad leaves was slightly decreased in bromeliads with predators relative to predator-free controls (`r predeffect("N")`%), and was only higher than the control in treatments including Tabanid predators.  We found a similar pattern for plant growth: on average, predators had a `r predeffect("growth")`% effect on growth of bromeliad leaves (leaf elongation in mm), though Tabanids seemed to create a slight increase. The decomposition of coarse detritus and production of fine organic matter showed no obvious pattern related to the presence of predators.

```{r polysummary}
polyeffect <- function(resp="total.surv"){
  diffeffect <- (mean(pd[[resp]][pd$treatment%in%c("elong + andro","elong + leech","elong + tab")],na.rm=TRUE)-mean(pd[[resp]][pd$treatment%in%c("andro","tabanid","leech","elong")],na.rm=TRUE))/mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE)
  round(diffeffect,digits=2)*100
  }

```

Predator combinations tended to have a non-additive effect on our response variables, even though we held total metabolic capacity constant amongst all treatments.  Approximately `r polyeffect()`% more prey survived in polyculture, on average,  compared to all monocultures.  Nitrogen uptake increased by (`r polyeffect("N")`%) and bromeliad growth by (`r polyeffect("growth")`%). Production of fine particulate organic matter increased by `r polyeffect("fine")`% more when predators were present in combination.

Our experimental design allows us to estimate the non-additive effect of predator species pairs on whole communities of prey, and the functioning of the bromeliad ecosystem.  We used randomization tests to test the hypothesis that increased phylogenetic distance between members of a predator pair results in a greater magnitude of nonadditive effect.  We contrasted the differences of the mean individual predator treatments from the control with the mean difference of their pairwise combination from the control.  

We found the greatest effect for prey survival: while effects of _L. andromache_ and _L. elongatum_ in combination were quite similar to the effect of either alone, when _L. elongatum_ was placed in the same plant as either a Tabanid larva or leeches, on average five more prey individuals (18% of total prey community) survived till the end of the experiment (Fig 3).  This effect was smaller among the other variables, most of which showed confidence intervals from the randomization test which overlap zero.


### Figures

```{r CALC_AIC_distribution and diet}
### Distribution
## calculates the AIC values that appear in the supp matt.
if(TRUE) {
  model_aic <- metabolic_occur_phylo %>% 
    ungroup %>% 
    select(phylodistance,overlap) %>%
    fit_some_models
  
  write.csv(model_aic,file = "./analyses/distributional_similarity.csv")
  }

### Diet

f_extract <- function(f,.models) sapply(.models,function(x) f(x[[1]]))

if(TRUE){
models <- diet_overlap_phylo %>% 
  ungroup %>% 
  select(phylodistance,overlap,nspp) %>%
  do(quadratic = nls(overlap ~ a * (phylodistance)^2 + b * phylodistance + c, 
                     data = ., 
                     start = list(a = -0.0000008, b = 0.0008, c = 1),
                     weights = nspp),
     bellshaped = nls(overlap ~ peak * exp(-1 * (phylodistance)^2 / a), 
                      data = ., 
                      start = list(a = 900000, peak = 1),
                      weights = nspp),    
     exponential = nls(overlap ~ b * exp(a * phylodistance),
                       data = .,
                       start = list(a = -0.0004, b = 1),
                       weights = nspp),
     #        Sshaped = nls(overlap ~ c * exp(a * phylodistance) / 
     #                        (c * exp(a * phylodistance) + (1 - c)),
     #                      data = .,
     #                      start = list(a = -0.007, c = 0.9)),
     linear = nls(overlap ~ a* phylodistance + b, 
                  data = .,
                  start = list(a = -0.1,b = 0.7), weights = nspp),
     constant = nls(overlap ~ z,
                    data = .,
                    start = list(z = 0.4),
                    weights = nspp)
     ) 


data.frame(model = names(models),
           AIC = f_extract(AIC,models)) %>%
  arrange(AIC) %>%
  write.csv(file = "./analyses/diet_similarity.csv")
}


```

```{r FIG_dist_diet_effect}
fig2 <- diet_overlap_phylo %>% 
  ungroup %>%
  select(phylodistance,overlap,nspp) %>%
  ggplot(aes(x = phylodistance, y = overlap)) + 
  geom_point(aes(size = nspp), shape  = 21) + 
  geom_point(aes(size = nspp), colour = "black", fill = "#00A08A", shape  = 21, alpha = 0.6) + 
  xlab("Phylogenetic distance") + ylab("Diet similarity (Pianka's index)") + 
  scale_size(range = c(3,9),name="Number of\nprey") +
  mytheme

# fig2 + stat_function(fun = exponential,args = f_extract(coef)$exponential %>% as.list , size = 1)

x_sequence <- diet_overlap_phylo  %>%
  ungroup %>%
  summarise(min_x = min(phylodistance),
            max_x = max(phylodistance)) %>%
  do(seq(from = .$min_x,
         to = .$max_x,
         length.out = 500) %>%
       data.frame(phylodistance = .)
     ) 

rawdata <- diet_overlap_phylo %>% 
  ungroup %>% 
  select(phylodistance,overlap,nspp)

predictions  <-  replicate(
  500,{
    boot  <- rawdata[sample.int(nrow(rawdata), replace = TRUE), ]
    #browser()
    model = nls(overlap ~ b * exp(a * phylodistance), 
                data = boot, 
                start = list(a = -0.0004, b = 1),
                weights = nspp)
    # Output predictions at each point that we'll want to plot later
    if(!is.null(model)) {
      predict(model, 
              data.frame(x = x_sequence))
      }
    else {
      rep(NA,length(x_sequence))
      }
    },
  simplify = FALSE
  )

predictions  <- do.call(cbind,predictions)

observed_fit <- nls(overlap ~ b * exp(a * phylodistance), 
                    data = rawdata, 
                    start = list(a = -0.0004, b = 1),
                    weights = nspp)
diet <- x_sequence %>% tbl_df %>%
  mutate(pred_m2 = predict(observed_fit,
                           newdata = list(phylodistance = phylodistance)),
         upper = apply(predictions,1,quantile,prob = .975, na.rm = TRUE),
         lower = apply(predictions,1,quantile,prob = .025, na.rm = TRUE)
         )

# %>% 
#   #  gather(model,prediction,-phylodistance) %>%
#   function(x) {fig2 + geom_line(aes(x = phylodistance, y = pred_m2),data = x) +
#      geom_line(aes(x = phylodistance, y = upper),data = x,linetype = "dashed") +
#      geom_line(aes(x = phylodistance, y = lower),data = x,linetype = "dashed")}
   

distribution <- metabolic_occur_phylo %>% 
  ungroup %>% 
  select(phylodistance,overlap) %>%
  mutate(category = "distribution", nspp = 2)

diet <- diet_overlap_phylo %>% 
  ungroup %>%
  select(phylodistance,overlap,nspp) %>%
  mutate(category = "diet")

rbind(distribution, diet) %>%
  ggplot(aes(x = phylodistance, y = overlap, size = nspp)) + 
  geom_point(colour = "black", fill = "#00A08A",
             shape  = 21, alpha = 0.6) +
  facet_grid(category ~ ., scales = "free_y") + 
  xlab("Phylogenetic distance") + 
  ylab("Distributional similarity (Pianka's index)") + 
  mytheme

```

**Figure 1**: Phylogenetic distance and predator co-occurrence.  Each point represents a pair of predator species.  We calculated total metabolic capacity for each predator species in each bromeliad, and then calculated niche overlap between two predators using Pianka's index. Solid line shows the best fit non-linear model (see Appendix) and dashed lines show bootstrap 95% quantiles. 

```{r FIG_feeding_trial_as_phylo, fig.cap=""}




```

**Figure 2**:  Diet overlap decreases with phylogenetic distance, based on 237 feeding trials with the 8 major predator taxa found in this system.  Diet similarity is calculated as Pianka's index of niche overlap. Solid line shows the best fit non-linear model (see Appendix) and dashed lines show bootstrap 95% quantiles.





```{r FIG_PD_experiment_nonadditive,fig.cap=""}

summarize_randoms_phylo %>% 
  filter(response_var == "survival") %>%
  ggplot(aes(x=phylodistance,y=mean)) + 
  geom_linerange(aes(ymin=lower, ymax=upper)) + 
  geom_point(size=6,shape = 21, colour = "black", fill = "#00A08A") + 
  ylab("Mean treatment difference, Control-Treatment") + 
  xlab("Time (Mya)") + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey") + 
  mytheme
# 
# 
# ggplot(summarize_randoms_phylo,
#        aes(x=Time,y=mean))+geom_errorbar(aes(ymin=lower, ymax=upper),width=0)+geom_point(size=3)+ylab("Mean treatment difference, Control-Treatment")+xlab("Time (Mya)")+facet_wrap(~variable)
```

**Figure 3**: Combinations of predators beyond congenerics show a negative non-additive effect on predation rate.  Points represent the mean difference between the means of two monocultures compared to the mean of a polycultures. The dashed line represents differences $= 0$, i.e. no nonadditive effect. These values are presented as differences from (predator free) controls; therefore zero indicates no significant non-additive effect. Relative to control (no predator) plants, bromeliads containing two predators which were not congeneric showed less predation


## Discussion

### Summary

We tested whether related predators are similar or dissimilar in their habitat distribution (Question 1).  We found only weak support: There is no evidence for close relatives being either more or less likely to be found in association.  Very distant relatives may be slightly likely to co-occur, as shown by the decelerating curve in Fig 1.

We also tested whether diet similarity decreased with predator phylogenetic distance (Question 2), and found a stronger pattern: diet similarity declined more rapidly than distributional similarity with PD, in a nearly linear relationship.

Question 3 has two parts. Question 3a hypothesized that differences among predator responses would correlate with their phylogenetic divergence, and we found no support for this pattern. We did find support for hypothesis 3b for one of our responses: increased phylogenetic diversity of the predator assemblage was associated with a negative non-additive effect on predation (increased prey survival).

### Do related organisms occur in different bromeliads? 

Predator pairs showed only a weak relationship between phylogenetic distance and differences in distribution.  

This may reflect a lack of variation among bromeliads, either in environmental variables or prey community composition, at least over the lifespans of the predators. 

It may also indicate that predator species in this system are habitat generalists.  Organisms which live in small habitats, which are prone to large fluctuations, may develop a wide physiological tolerance as they integrate over variation in prey community composition, temperature, etc.

We observed little predator mortality in our experiment; this indicates that perhaps even negative interactions among predators would not necessarily lead to differences in their distribution.

The trend in Figure 1 is largely caused by the distribution of leeches among bromeliads. The leeches are both distantly related to the other predator taxa (all holometabolous insect species) and most different from them in distribution. It is not clear why this should be, but we can speculate that we are seeing the result of different dispersal modes: of all the large predators, only leeches disperse passively.

This indicates that a range of phylogenetic diversity is present in nature, with some bromeliads containing close relatives and others more distant. It also supports our experimental design by suggesting that our predator combinations -- even the most disparate -- are common in nature.

There are several potential reasons why the negative intraspecific effects that we observed in our experiment do not appear to limit predator distribution. First of all, the probability of interaction in real bromeliads could be quite low, especially in larger bromeliads than those used in our experiment.  The negative effects could still be occurring, but not resulting in mortality. Indeed, if animals are actually able to adjust their development times and feeding rates when conditions are unfavourable (and if the presence of a competing predator creates such unfavourable conditions), then we would expect to observe more frequent, not less frequent, co-occurances of predators.  Negative interactions need not result in dissimilar distributions or "checkerboard units".

### Do related organisms have similar diets?

Predators consumed very similar prey, except at the most extreme differences in the phylogeny.

Leeches and tabanids showed different patterns of consumption compared with *Leptagrion* predators: they consumed less frequently, and notably consumed prey in less than half of trials with *Culex* (leech) or *Scirtes* (Tabanid).  This may be because these predators lack the strong jaws and "mask" of odonata, and these two prey species are particularly active (in the case of culicids) or difficult to handle (scirtids).  Tabanids in particular differ in microhabitat use from odonates; living deep within leaf axils, where culicids are rarely found.

One of our morphospecies, *Leptagrion* "tan", showed a greater preference for harder-bodied prey species (i.e. Ostracoda, Scirtes and *Phylloicus*, which is a caddisfly.)  All of these animals are consumed more frequently by this animal than by the other species of Leptagrion, while the other damselflies more frequently consumed softer *Culex* and *Polypedilum* prey.  

Traits can be more important than phylogeny *per se* to a predator's diet: [@Moody1993] found that unrelated decapod species which were morphologically similar were also functionally similar.
Similarly, [@Rezende2009] found that both body size and phylogeny determined the food web "compartment" (shared predator-prey interactions) of a predator in a marine foodweb.

Our experimental results are consistent with high similarity among predator diets: all predator species had comparable effects on all response variables, including prey species survival.

### Phylogenetic diversity and non-additive effects

Predator pairs which were more phylogenetically distant had a larger negative nonadditive effect on prey capture.  This is contrary to one of our first predictions, which suggested that more distant predators would show an increase in prey capture.  

It is intriguing that the congener *L. andromache* did not produce a trait-mediated effect in *L. elongatum*, while more phylogenetically distinct predators did. It may be that these odonates have behavioural traits that reduce the probability of their interaction, for example each nymph occupying a single leaf-well (Srivastava, CR).  If this is the case, each damselfly may not experience many cues indicating the presence of other predators, resulting in no non-additive interaction.

Odonates can be sensitive to chemical cues of potential predators, which causes a decrease in feeding rate (Barry and Roberts 2014). However other research in bromeliads has demonstrated that it is physical contact with other organisms that reduces damselfly predation rate (Trish).  If this is the case for *Leptagrion*, then when combined with leeches they may be responding to frequent contact with those very active predators. Tabanids, however, are rarely observed outside of a deep leaf axil -- in this case, it may be chemical cues which are responsible. 

In feeding trials with leeches, actual predation was rarely observed -- except in one instance, where the damselfly ate the leech. However, trait-mediated indirect effects may not always reflect realized probability of predation: an animal may change behaviour when exposed to a "predator" poses little threat (REF).

One limitation of our approach is the focus on a single focal predator, the odonate *Leptagrion elongatum*, which was common in all of our predator treatments. It is possible that this species is more sensitive to the presence of other predators, and therefore shows a larger trait-mediated indirect effect, than would other species in this community.  However, this is the most common species in this community and our results indicate that its top-down effects are likely to be frequently reduced by the presence of other predators.

## References