---
title: "Predator phylogenetic diversity decreases predation rate via antagonistic interactions"
author: "A. Andrew M. MacDonald; Gustavo Q. Romero; Diane S. Srivastava"
fontsize: 12pt
output:
  pdf_document:
    includes:
      in_header: formatting/myheader.tex
    latex_engine: xelatex
  md_document: default
  html_document: default
  word_document:
    reference_docx: formatting/reference.docx
geometry: margin=1in
bibliography: /home/andrew/Documents/reference/Mendeley_reference_lists/@MS_pdpaper.bib
---

```{r options_packages_data, message=FALSE, echo=FALSE}
## set chunk options
library(knitr)
opts_chunk$set(message=FALSE,warning=FALSE,echo=FALSE, fig.cap=TRUE,
               cache = TRUE, cache.path = "./cache/",
               dev='pdf', fig.path = './figure/')

## load packages
library(ggplot2)
library(bipartite)
library(vegan)
library(picante)
library(pander)
library(dplyr)
library(magrittr)
library(xtable)
library(stargazer)
library(tidyr)
library(reshape2)

## ggplot2 theme
mytheme <- theme_bw()

# occurrance data -- abundance
occur <- read.csv("../data/predator.cooccur.txt",
                  stringsAsFactor=FALSE) %>%  tbl_df()
# occurrance data -- metabolic capacity
metabolic  <- read.csv("../data/predator.cooccur.metabolic.txt",
                       stringsAsFactor=FALSE) %>%  tbl_df()

metabolic_occur_phylo <- read.csv("../data/metabolic_occur_phylo.csv",
                                  stringsAsFactors = FALSE)

# feeding trial data
feeding_trials <- read.csv("../data/reorganized.feeding.trial.data.csv",
                    stringsAsFactors=FALSE) %>%   tbl_df()


diet_overlap_phylo <- read.csv("../data/diet_overlap_phylo.csv",
                    stringsAsFactors=FALSE) %>%   tbl_df()

# experimental data
pd <- read.csv("../data/pd_exp_cleaned_data.csv",
               stringsAsFactors=FALSE) %>%   tbl_df()

summarize_randoms_phylo <- read.csv("../data/summarize_randoms_phylo.csv")


# enriched leaves -- ie just the N data for the detritus we put in.
# enriched <- read.csv("../data/enriched_leaves.csv",
#                      stringsAsFactors=FALSE) %>%   tbl_df()
# phylogeny data
nodeages <- read.csv("../data/nodeages.csv")
  

# randomization tests
summarize_random_test <- read.csv("../data/randomization_summary.csv")

## load in functions
source("../R.scripts/FUNCTIONS_predator_diversity.R")

source("../R.scripts/FUNCTIONS_text.R")
```

```{r experiment_data_postprocess}
#dropping a record that seems to have been 90% decomposed!
pd <- pd %>%
  transform(decomp=ifelse(decomp>0.7,NA,decomp))
```

## Introduction

Predators can have strong top-down effects, both on community structure and
ecosystem processes [@Estes2011]; however their effects in combination are not
well understood. The net effect of a predator assemblage is often dependent on
species identity, and can be more or less than the additive effects of
individual predators [@Sih1998a;@Ives2005]. Non-additive effects of predator
diversity occur when predators interact directly with each other, or indirectly
via prey species, in a way that increases or decreases total predation.
Antagonistic among-predator interactions can reduce the top-down effects of a
diverse consumer assemblage, for example when predators feed directly on each
other (intra-guild predation), consume the same prey (resource competition) or
modify the behaviour of predators or prey via non- consumptive (i.e. trait-
mediated) interactions [@Sih1998a;@Griswold2006;@Nystrom2001]. Synergy between
predators is also possible, for example if one predator causes behavioral shifts
in prey that increase their chance of being eaten by another predator species.
These interactions depend on predators actually co-occurring, and predators in
the same community may not always co-occur, for example because of microhabitat
preference differences [@Schmitz2007]. While there are many mechanisms
potentially underlying predator compositional effects, we lack a means of
predicting _a priori_ which combinations are likely to have antagonistic versus
synergistic effects on ecosystem functions.

One possible predictor of the effect of predator combinations is predator
phylogenetic diversity. The measurement of phylogenetic diversity has become a
popular means for ecologists to make inferences about ecological and
evolutionary mechanisms behind observed patterns in natural communities
[@Cavender-Bares2009]. For example, the phylogenetic diversity of plant
communities is a better predictor of productivity than is either species
richness or diversity [e.g. @Cadotte2009;@Cadotte2008;@Godoy2014]. A central
assumption of this approach is that increased phylogenetic distance implies
increased ecological dissimilarity -- either in the form of differences in
species niches, interactions, or functional traits. When this is true, high
phylogenetic diversity could creat the ecological conditions neede for strong
ecosystem functioning [@Srivastava2012c]. Despite the prevalence of phylogenetic
community ecology and the importance of predators to natural systems, the
phylogenetic diversity of local predator assemblages has rarely been measured
[@Bersier2008;@Naisbit2011]. Many studies of phylogenetic signal in predator
traits focus on whole clades, rather than local assemblages (e.g. *Anolis*
lizards [@Knouft2006], warblers [@Bohning- Gaese2003], tree boas
[@Henderson2013] and wasps [@Udriene2005]) making it difficult to connect these
results to top- down effects at the scale of a local community. While these
clade-specific studies often find weak evidence for phylogenetic signal in
ecologically- relevant traits, studies at the level of the whole biosphere
[@Gomez2010;@Bersier2008] demonstrate that related organisms often have similar
interspecific interactions -- i.e. related predators often consume similar prey.
At the level of a community, the effect of a predator assemblage will depend on
both the microhabitat distribution and trophic interactions of all predator
species. Phylogeny may constrain local species composition when distant
relatives have distinct fundamental niches, while close relatives are too
similar to co-exist [@Webb2002;@Emerson2008]. Other predators constrain
distribution not only through direct interactions, but also colonization:
predators may use cues of prey, competitors, or intra-guild predators when
selecting a habitat patch [@Chase2009a;@McCauley2008a;@Srivastava2008]. After
colonization, patterns of co-occurrence may be further modified by intraguild
predation [@Lounibos2008;@Juliano2009] or competition [@Fincke1997].

When predators do co-occur, their direct effects on each other and on their prey
is in part determined by their overlap in diet. Diet overlap (shared prey
species between predators) will depend on the feeding traits and nutritional
requirements of predators -- both of which may be phylogenetically conserved
(CITEAngelica). If this is the case, then predator assemblages with higher
phylogenetic diversity will show greater prey consumption and a stronger top-
down effect on ecosystem function [@Finke2008a]. For example, some predators are
gape-limited, and can swallow any prey smaller than their mouth; these predators
typically broaden their diet as they grow (e.g., larvae of Odonata)
[@Werner1984].  Such ontogenetic changes in diet can radically influence
food web structure [@Moya-Larano2011]. Other predators (e.g. Leeches, or
Diptera:Tabanidae) are "piercing-sucking" predators [@Bay1974] and may be more
general consumers at all life stages, since they can feed on larger prey
individuals even at very early predator life stages.  In some cases, predator
diets may extend to include other predators, leading to direct negative
interactions such as intraguild predation consequently have a phylogenetic
signal [@Pfennig2000].

<! -- 
Indirect interactions among predators are common in many communities, which
means that the total amount of top-down control cannot be predicted by summing
the contributions of individual predator species [@Sih1998a]. For example,
decreasing predator richness has been shown to increase herbivory in a three-
level kelp food web [@Byrnes2006]. The complex traits that underlie such
species interactions ("higher-order" interactions, Sih *et al.* [-@Sih1998a])
may also correlate with phylogeny; if so, then phylogenetic diversity will be
a better predictor of ecosystem function than richness alone [@Cadotte2009].
The effect of predators on prey may also be non-additive, the result of
indirect interactions either with prey species or other predators: prey
species may respond to the presence of one predator by a behavioural shift
that increases the predation by another predator population [@Carey2010a].
These trait-mediated indirect effects (TMII, [@Werner2003]) can be difficult
to predict; however phylogenetic relatedness has been suggested as a possible
means of predicting the combined effect of predators [@Naisbit2011]
-->

We used a series of observations, lab feeding trials, and manipulative field
experiments to measure how the phylogenetic diversity of the predator
assemblage predicts community composition and ecosystem function.  We test
three related hypotheses concerning similarity in distribution, diet and top-
down ecosystem effects of predators, using a natural mesocosm: the community
of invertebrates living within bromeliads. Bromeliads (Bromeliaceae) are
epiphytic plants native to the Neotropics. Many bromeliad species contain
water and detritus; the decomposition of this detritus supplies nutrients for
the bromeliad [@Reich2003]. The small size of these habitats permits direct
manipulations of entire food webs, manipulations which would be difficult in
most natural systems.  Within this aquatic food web, damselfly larvae (e.g.
*Leptagrion* spp., Odonata:Coenagrionidae) are important predators; their
presence dramatically alters community dynamics (e.g. decreases insect
emergence [@Starzomski2010] and increases nutrient cycling [@Ngai2006]).

1. *species co-occurrence*: closely-related predators may occur together more
frequently than less-related predators if there is a strong phylogenetic
signal to habitat requirements.   Alternatively, very closely related species
may never co-occur because high overlap in ecological niches results in
competitive exclusion.

2. *diet similarity*: similarity in diet (as measured by feeding trials)
decreases with phylogenetic distance if diet is phylogenetically conserved.
Alternatively,  closely related species may have evolved different diets to
allow coexistence.

3. *ecosystem-level effects*: Our experiments at the level of the whole habitat patch (i.e. a single bromeliad) allow us to test hypotheses about direct and indirect effects of predator combinations:
    (a) Closely-related predators will have similar effects on the community. This will occur if related predators share similar trophic interactions (e.g. predation rate, diet similarity). Our monoculture treatments allow us to assess the effect of each predator on both a complete prey community and ecosystem function.  
    (b) More phylogenetically diverse predator assemblages will show a stronger non-additive effect. This will occur if phylogenetic distance correlates with increasing trait difference, and if this trait difference is enough to lead to complementarity. At the extreme, differences between predators may lead to IGP among predators. By comparing treatments with pairs of predators to treatments which received a monoculture of each predator, we are able to estimate additive and nonadditive effects. 

## Methods

### Site and Species

We conducted all observations and experiments in Parque Estadual da Ilha do
Cardoso ($25^{\circ} 03^{\prime}$ S, $47^{\circ}53^{\prime}$ W), a 22.5 ha
island off the south coast of São Paulo state, Brazil. We worked in a closed
coastal forest (*restinga*) the understory of which is mostly covered by
*Quesnelia arvensis* Mez. (Bromeliaceae), a large terrestrial bromeliad that
accumulates up to 2.8 L of rainwater in tanks formed by individual leaves. We used two datasets: an observational survey of insect diversity from 2008, and a manipulative experiment in 2010. Our observational survey found more than 47 species of macroinvertebrates in these aquatic communities [@Romero2010].
This diversity encompasses multiple trophic and functional groups: filter feeders (Diptera:Culicidae); detritivores including shredders
(Diptera:Tipulidae, Trichoptera), scrapers (Coleoptera:Scirtidae), and collectors
(All Diptera:Chironomidae, Syrphidae, Psychodidae). These species are the main prey for a 
diverse predator assemblage of at least three species of damselfly larvae (_Leptagrion_ spp., Zygoptera:Coenagrionidae), two species of predatory Horse Fly larvae (Diptera:Tabanidae), and 2
species of leech (Hirudinidae).  Other, smaller-bodied predators present include predatory midge larvae (Diptera:Chironomidae:Tanypodinae) and less abundant species such as Dytiscid beetles.

### Distributional similarity: observational data

We investigated Question 1 above with observations of predator distribution
among bromeliads, using a 2008 survey of bromeliad communities.  Each bromeliad
was dissected and washed to remove invertebrates and the resulting water was
filtered through two sieves (250 and 850 µm), which removed particulate organic
matter without losing any invertebrates. All invertebrates were counted and
identified to the lowest taxonomic level possible. The body length of all
individuals was measured, when possible for small and medium-sized taxa (< 1cm
final instar) and always for large-bodied taxa (> 1 cm final instar).

### Diet Similarity: feeding trials
```{r CALC_feeding_trials}
## numbers for text
ntrials <- sum(feeding_trials$number.trials)
npred <- length(unique(feeding_trials$predator.names))
nprey <- length(unique(feeding_trials$Prey.species))
ncombos <- nrow(feeding_trials)
rep_range <- range(feeding_trials$number.trials)
```

We tested for similarity in predator diets (Question 2) using a series of laboratory feeding trials. We conducted `r ntrials` feeding trials of `r npred` predator taxa fed `r nprey` prey taxa between March and April 2011.  We  covered all potential predator-prey pairs present in the experiment
(described below), and attempted to perform all other trials whenever possible. However, due to the rarity of some taxa many predator-prey pairs were not possible; we tested `r ncombos` pairwise combinations.  Most trials were replicated at least five times, but the number of replicates ranged from `r rep_range[[1]]` to `r rep_range[[2]]`. We placed
predators together with prey in a 50ml vial, with a leaf or stick for
substrate. The only exception was the tabanid larvae, which we
placed between two vertical surfaces to imitate the narrow space found in
bromeliad leaf axils (their preferred microhabitat).  Generally our trials contained a single predator and a
single prey individual, except in the case of very small prey (_Elpidium_ sp.)
or predators (_Monopelopia_ sp.) in which case we increased the density.  We
replicated each combination up to five times where possible, and allowed 24 hours for
predation to occur.

### Community effect experiment

Our third question had two parts: (a) how do predator species differ in their
effects on the whole community and (b) do predators show nonadditive effects on
the community in combination, and are these related to phylogenetic distance? We
tested these questions with an experiment involving the four most common
predators, paired to create a range of phylogenetic relatedness: two congeneric
damselflies (*Leptagrion andromache* and *Leptagrion elongatum*), two insects
(*L. elongatum* and a Tabanid predatory fly), and two invertebrates (*L.
elongatum* and a predatory leech).  Our experiment also included species in
monoculture, and a predator-free control (total of 8 treatments, n=5).  Treatments were
substitutive, maintaining the same amount of total predator metabolic capacity
(see below) to isolate the effects of predator behaviour and traits.
Substitutive experiments often hold total abundance constant, but when species
differ substantially in body size - as in this experiment - allometric effects
of body size on feeding rate can confound detection of effects based on trophic
traits or species interactions, and standardizing to community metabolic
capacity is preferred [@Srivastava2009a]. This experiment allows the estimation
of the effect of each predator species (monoculture treatments), as well as the
detection of non- additive effects in predator combinations.

In Feburary 2011, bromeliads between 90 and 200ml in capacity were collected,
thoroughly washed to remove organisms and detritus and soaked for 12 hours in a
tub of water. They were then hung for 48 hours to dry.  One bromeliad dissected
after this procedure contained no insects. Each bromeliad was supplied with
dried leaves, simulating natural detritus inputs from the canopy. In order to
track the effects on detrital decomposition on bromeliad nutrition, we enriched
these leaves with ^15^N by fertilizing five (Jabuticaba, *Plinia cauliflora*)
plants with 40ml pot^-1^ day^-1^ of 5g L^-1^ ammonium sulphate containing 10%
atom excess of ^15^N over 21 days. Whole leaves were then picked from plants and
air-dried until constant weight, and then soaked for three days and the water
discarded. About 1.5 g of leaves were placed in each bromeliad (`r mean(pd$mass.g.)`g ± `r sd(pd$mass.g.)`).

Each bromeliad was stocked with a representative insect community. The densities
of each prey taxon were calculated from the 2008 observational dataset, using
data from bromeliads of similar size to those in our experiment.  All densities
used were within the range of these calculated abundances, and all experimental
bromeliads received the same insect community.  Halfway through the experiment,
insects were added to bromeliads a second time to simulate the continuous
oviposition that characterizes the system.  Throughout the experiment, all bromeliads were enclosed
with a mesh cage topped with a malaise trap and checked daily for emergence of
adults.

In this experiment we measured five response variables: production of fine
particulate organic matter (FPOM), decomposition of coarse detritus, bromeliad
growth, uptake of detrital nitrogen into bromeliad tissue, and survival of invertebrate prey (emerged adults +
surviving larvae). We analyzed each of these responses with ANOVA.  For each pair of predator species and each response type, we
calculated the non-additive effect as the difference between the
response in bromeliads with both predator species (n=5) and the mean response in bromeliads with either one of these two predator species 
(n=5 for each predator species).  We generated bootstrap confidence intervals for
these nonadditive effects; confidence intervals which do not overlap zero
indicate a significant non-additive effect of a predator combination.

### Phylogenetic distance, metabolic capacity and niche overlap

```{r CALC_nodeages}
n.nodes <- nrow(nodeages)
nstudies <- sum(nodeages$nstudies)
```
Our observational data contains data from `r ncol(metabolic)-1` bromeliads, containing 
`r metabolic$Taxa %>% unique %>% length` predator taxa. These include two groups of 
congeneric taxa -- _Bezzia_ sp. (Diptera:Ceratopogonidae) with two species and _Leptagrion_ sp.
(Odonata:Coenagrionidae) with three species. There were also two groups of confamilial taxa -- three species of
Tabanidae and two species of Empididae, all Diptera.  Deeper divisions were also present: three families of Diptera are represented by a single predator species each (Dolichopodidae, Corethrellidae and
Chironomidae) and the deepest taxonomic divide is between all insects present and the predatory leeches (Annelida:Hirudinidae).

We obtained node age estimates for all seven internal nodes of the tree, using `www.timetree.org`, an online database of published  molecular time estimates [@Hedges2006].  Most dates came from only a single study; where multiple dates were found we used the median estimate: Insecta--Hirudina (`r agerange("insects.to.leeches")` Mya, n=`r nstudy("insects.to.leeches")` studies), Odonata--Tabanidae (`r agerange("odonata-Tabanidae")` Mya, n=`r nstudy("odonata-Tabanidae.csv")` studies) and Tabanidae--Diptera (`r agerange("tabanidae_culidicae_ie_Diptera")` Mya, n=`r nstudy("tabanidae_culidicae_ie_Diptera")` studies).  Node age data was available
for all but the youngest nodes, where either a lack of taxonomic
information (e.g. Tabanidae) or a lack of phylogenetic study (e.g. _Leptagrion_)
prevented more information from being included.   These branches were left as
polytomies, and were all assigned identical, arbitrary and short branch lengths
(15 Mya).

Species co-occurrence is often measured in terms of non-random patterns of
species presence/absence or abundance, but such measures will only be poorly related
to the functional effects of species when species differ substantially in body
size.  Integrating the allometric relationship between body size and feeding
rate [@Brown2004; @Wilby2005] over all individuals of a species allows estimates
of "metabolic capacity", or the potential energy requirements of a species
[@Srivastava2009a]. Metabolic capacity is equal to individual body
mass raised to the power of 0.69 (an insect-specific exponent determined by Peters 19xx and confirmed by Chown et al, [-@Chown2007]); this reflects the nonlinearity of feeding rate
on body size across many invertebrate taxa. We used metabolic capacity to inform
both our observational results and our experimental design (details below), with the exception
of our feeding trial data.  This is because the feeding trials were intended to
measure which prey our predators ate, rather than their feeding rate (only the
latter should scale with metabolic capacity).

Questions 1 and 2 above focus on similarity ("niche overlap") between predator species. Question 1 hypothesizes that if habitat preference is phylogenetically conserved and competition between predators unimportant, then distributional similarity will decline with increased phylogenetic distance between a pair of predators. Question 2 hypothesizes that if diet is conserved, then diet similarity declines with increased phylogenetic distance between a pair of predators. We evaluated both distributional and diet similarity between predators using Pianka's index of niche overlap [@Pianka1974]:

$O_{kl}=\dfrac{\sum_i^n{p_{il} p_{ik}}}{\sqrt{\sum_i^n{p_{il}^2} \sum_i^n{p_{ik}^2}}}$

For each pair of predators, $p_{ik}$ and $p_{il}$ represent the preference of
predator $k$ (or $l$) for resource or habitat $i$. When quantifing distributional similarity, $n$=`r ncol(metabolic)-1` bromeliads surveyed in the observational data; when quantifying diet similarity, the number of resources ($n$) is defined as the total number of prey
species assayed with both predator taxa.  In the survey data, preference ($p_{ik}$) is defined as the proportion of a predator's total metabolic capacity found in a particular bromeliad; in diet trials it is the
proportion of diet trials (see below) for each predator-prey combination that resulted in prey
mortality.

We quantified the effect of phylogenetic distance on each of distributional and diet similarity.  First, we calculated phylogenetic distance between each pair of species, then fit several functions to the relationship between phylogenetic diversity an either distributional or diet similarity (linear, constant, and several appropriate nonlinear functions).  We compared these models using AIC, selected the best model, and generated confidence intervals as appropriate (parametric or bootstrap for linear and nonlinear, respectively).

## Results

### Distributional similarity
```{r CALC_metabolic_numbers}
## numbers for text:
meta_range <- metabolic %>%
  melt(id.vars="Taxa") %>%
  select(-variable) %>%
  filter(value>0) %>%
  filter(value%in%range(value)) %>%
  transform(value=signif(value,2)) %>%
  unique()

smallsp <- meta_range[1,]
largesp <- meta_range[2,]

nspp <- metabolic %>%
  melt(id.vars="Taxa",variable.name="brom") %>%
  filter(value>0) %>% 
  group_by(brom) %>%
  summarise(nspp=n()) %>%
  summarise(meanpred=mean(nspp),
            sdpred=signif(sd(nspp),2)) 

```

```{r TEST_metabolic~phylo}
meta_phylo_lm_summary <- lm(overlap~phylodistance, data = metabolic_occur_phylo) %>% summary
```

Across all bromeliads, predator species differed widely in metabolic capacity, from 4.5 x 10^-4^ for a species of _Monopelopia_ to `r largesp$value` for large predatory flies (Tabanidae).  Predators often co-occurred in bromeliads (`r nspp$meanpred` ± `r nspp$sdpred` predator species per plant).  However, the niche overlap between the total metabolic capacity of pairs of predators did not show any relationship with phylogenetic distance between them (`r F_text(meta_phylo_lm_summary)`).

### Diet similarity

```{r CALC_diet_numbers}
## Diet numbers for text:
## what percentage of total trials resulted in predation?
percentpredation <- feeding_trials %>%
  transform(percent.eaten = eaten/number.trials) %>%
  group_by(predator.names) %>%
  summarise(mean.pred = mean(percent.eaten),
            totaltrials = sum(number.trials),
            nprey = n())
  
andro <- percentpredation %>%
  filter(predator.names == "Leptagrion.andromache") %>%
  select(mean.pred) %>%
  transform(mean.pred=round(mean.pred*100,1)) %>%
  as.numeric()

elong <- percentpredation %>%
  filter(predator.names=="Leptagrion.elongatum") %>%
  select(mean.pred) %>%
  transform(mean.pred=round(mean.pred*100,1)) %>%
  as.numeric()
```

```{r TEST_diet-sim_as_PD,results='asis'}
diet_phylo_lm_summary <- diet_overlap_phylo %>%
  #filter(nspp>1) %>%
  lm(overlap~phylodistance, weights = nspp, data = .) %>%
  summary()

dietmat <- diet_overlap_phylo %>%
  select(-species_pair) %>%
  unique() %>%
  dcast(phylopred1~phylopred2,value.var="overlap") %>%
  data.frame(row.names="phylopred1") %>%
  as.matrix()
```


Among the most common predator taxa (i.e. those used in our experiment, described below) the damselflies (_Leptagrion andromache_ and _Leptagrion elongatum_) showed the highest rates of prey consumption (prey consumed in `r andro`% and `r elong`% of trials, respectively). All predators showed a broadly similar diets, consuming nearly all species offered to them.  However, more phylogenetically distant predators differed in their preference of prey species, as measured by the niche overlap index (`r F_text(diet_phylo_lm_summary)`), regression weighted by the number of prey species assayed.)

### Ecosystem-level effects and phylogenetic distance

```{r experimentsummary}
predeffect <- function(resp="total.surv"){
  diffeffect <- (mean(pd[[resp]][pd$treatment!="control"],na.rm=TRUE)-mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE))/mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE)
  round(diffeffect,digits=2)*100
  }
```

Predators had a large effect on prey survivorship: on average all predator treatments showed `r abs(predeffect())`% lower prey emerging or surviving as larvae relative to the predator-free control.  Nitrogen transport to bromeliad leaves was slightly decreased in bromeliads with predators relative to predator-free controls (`r predeffect("N")`%), and was only higher than the control in treatments including Tabanid predators.  We found a similar pattern for plant growth: on average, predators had a `r predeffect("growth")`% effect on growth of bromeliad leaves (leaf elongation in mm), though Tabanids seemed to create a slight increase. The decomposition of coarse detritus and production of fine organic matter showed no obvious pattern related to the presence of predators.

```{r polysummary}
polyeffect <- function(resp="total.surv"){
  diffeffect <- (mean(pd[[resp]][pd$treatment%in%c("elong + andro","elong + leech","elong + tab")],na.rm=TRUE)-mean(pd[[resp]][pd$treatment%in%c("andro","tabanid","leech","elong")],na.rm=TRUE))/mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE)
  round(diffeffect,digits=2)*100
  }

```

Predator combinations tended to have a non-additive effect on our response variables.  Approximately `r polyeffect()`% more prey survived in polyculture, on average,  compared to all monocultures.  Nitrogen uptake increased by (`r polyeffect("N")`%) and bromeliad growth by (`r polyeffect("growth")`%). Production of fine particulate organic matter increased by `r polyeffect("fine")`% more when predators were present in combination.

We tested the hypothesis that increased phylogenetic distance between members of a predator pair results in a greater magnitude of nonadditive effect using randomization tests.  We contrasted the differences of the mean individual predator treatments from the control with the mean difference of their pairwise combination from the control. We found the greatest effect for prey survival: while effects of _L. andromache_ and _L. elongatum_ in combination were quite similar to the effect of either alone, when _L. elongatum_ was placed in the same plant as either a Tabanid larva or leeches, on average five more prey individuals (18% of total prey community) survived till the end of the experiment (Fig 3).  This effect was smaller among the other variables, most of which showed confidence intervals from the randomization test which overlap zero.


### Figures

```{r CALC_AIC_distribution and diet}


### Diet

f_extract <- function(f,.models) sapply(.models,function(x) f(x[[1]]))

diet_models <- diet_overlap_phylo %>% 
  ungroup %>% 
  select(phylodistance,overlap,nspp) %>%
  do(quadratic = nls(overlap ~ a * (phylodistance)^2 + b * phylodistance + c, 
                     data = ., 
                     start = list(a = -0.0000008, b = 0.0008, c = 1),
                     weights = nspp),
     bellshaped = nls(overlap ~ peak * exp(-1 * (phylodistance)^2 / a), 
                      data = ., 
                      start = list(a = 900000, peak = 1),
                      weights = nspp),    
     exponential = nls(overlap ~ b * exp(a * phylodistance),
                       data = .,
                       start = list(a = -0.0004, b = 1),
                       weights = nspp),
     #        Sshaped = nls(overlap ~ c * exp(a * phylodistance) / 
     #                        (c * exp(a * phylodistance) + (1 - c)),
     #                      data = .,
     #                      start = list(a = -0.007, c = 0.9)),
     linear = nls(overlap ~ a* phylodistance + b, 
                  data = .,
                  start = list(a = -0.1,b = 0.7), weights = nspp),
     constant = nls(overlap ~ z,
                    data = .,
                    start = list(z = 0.4),
                    weights = nspp)
     ) 


data.frame(model = names(diet_models),
           AIC = f_extract(AIC,diet_models)) %>%
  arrange(AIC) %>%
  write.csv(file = "./analyses/diet_similarity.csv")



```


```{r nonlinear_trendline_confint}

x_sequence <- diet_overlap_phylo  %>%
  ungroup %>%
  summarise(min_x = min(phylodistance),
            max_x = max(phylodistance)) %>%
  do(seq(from = .$min_x,
         to = .$max_x,
         length.out = 500) %>%
       data.frame(phylodistance = .)
     ) 

rawdata <- diet_overlap_phylo %>% 
  ungroup %>% 
  select(phylodistance,overlap,nspp)

predictions  <-  replicate(
  500,{
    boot  <- rawdata[sample.int(nrow(rawdata), replace = TRUE), ]
    #browser()
    model = nls(overlap ~ b * exp(a * phylodistance), 
                data = boot, 
                start = list(a = -0.0004, b = 1),
                weights = nspp)
    # Output predictions at each point that we'll want to plot later
    if(!is.null(model)) {
      predict(model, 
              data.frame(x = x_sequence))
      }
    else {
      rep(NA,length(x_sequence))
      }
    },
  simplify = FALSE
  )

predictions  <- do.call(cbind,predictions)

observed_fit <- nls(overlap ~ b * exp(a * phylodistance), 
                    data = rawdata, 
                    start = list(a = -0.0004, b = 1),
                    weights = nspp)

diet_predictions <- x_sequence %>% tbl_df %>%
  mutate(pred_m2 = predict(observed_fit,
                           newdata = list(phylodistance = phylodistance)),
         upper = apply(predictions,1,quantile,prob = .975, na.rm = TRUE),
         lower = apply(predictions,1,quantile,prob = .025, na.rm = TRUE),
         category = "(b)",
         category = factor(category,levels = c("(b)", "Distribution"))
         )
```


```{r FIG_dist_diet_effect, fig.height=9}

distribution <- metabolic_occur_phylo %>% 
  ungroup %>% 
  select(phylodistance,overlap) %>%
  mutate(category = "(a)", nspp = 2)

diet <- diet_overlap_phylo %>% 
  ungroup %>%
  select(phylodistance,overlap,nspp) %>%
  mutate(category = "(b)") %>%
  as.data.frame

exper <- pd %>%
  filter(treatment %in% c("andro","elong","tabanid","leech")) %>%
  select(treatment, Culicidae:Scirtidae) %>%
  group_by(treatment) %>%
  summarise_each(funs(mean)) %>%
  paired_predator_pianka(pred_colname = "treatment", ... = -treatment) %>%
  ungroup %>%
  mutate(species_pair = ifelse(species_pair == "andro_elong",
                               "Leptagrion.andromache_Leptagrion.elongatum", species_pair),
         species_pair = ifelse(species_pair == "andro_leech",
                               "Leptagrion.andromache_Hirudinidae", species_pair),
         species_pair = ifelse(species_pair == "andro_tabanid",
                               "Leptagrion.andromache_Tabanidae.spA", species_pair),
         species_pair = ifelse(species_pair == "elong_leech",
                               "Leptagrion.elongatum_Hirudinidae", species_pair),
         species_pair = ifelse(species_pair == "elong_tabanid",
                               "Leptagrion.elongatum_Tabanidae.spA", species_pair),
         species_pair = ifelse(species_pair == "leech_tabanid",
                               "Tabanidae.spA_Hirudinidae", species_pair)) %>%
  left_join(phylogenetic_distance) %>% 
  ungroup %>%
  select(phylodistance,overlap,nspp) %>%
  mutate(category = "(c)")

rbind(distribution, diet, exper) %>%
  ggplot(aes(x = phylodistance, y = overlap, size = nspp)) + 
  geom_point(colour = "black", fill = "#00A08A",
             shape  = 21, alpha = 0.6) +
  facet_wrap( ~ category, scales = "free_y", nrow = 3) + 
  xlab("Phylogenetic distance") + 
  ylab("Similarity (Pianka's index)") + 
  scale_size(range = c(3,9),name = "Number of\nprey") +
  geom_line(aes(x = phylodistance, y = pred_m2), 
            size = 0.5, data = diet_predictions) +
  geom_line(aes(x = phylodistance, y = upper),
            size = 0.5, data = diet_predictions, linetype = "dashed") +
  geom_line(aes(x = phylodistance, y = lower),
            size = 0.5, data = diet_predictions, linetype = "dashed") + 
  mytheme + theme(strip.text = element_text(hjust = 0.01),
                  strip.background = element_blank())

```

**Figure 1**: Phylogenetic distance between predator pairs and their ecology.  Each point represents a pair of predator species. Distributional similarity was quantified from the total metabolic capacity for each predator species in each bromeliad. Diet similarity is calculated from a series of feeding trials, and experimental similarity as the surviving prey community emerging from bromeliads. Similarity between each pair of predators was calculated using Pianka's index. Solid line shows the best fit non-linear model (see Appendix) and dashed lines show bootstrap 95% quantiles. 

```{r FIG_predator_pres_number, fig.cap=""}

# 
pred_present <- pd %>%
  select(treatment, total.surv) %>%
  group_by(treatment) %>%
  summarise(trtmeans = mean(total.surv)) %>%
  mutate(ispred = ifelse(treatment == "control", "absent", "present")) %>%
  group_by(ispred) %>%
  mutate(presence_mean = mean(trtmeans))
  
pres <- pred_present %>%
  ggplot(aes(x = ispred, y = trtmeans)) + 
  geom_point(colour = "black", fill = NA, shape = 21, colour = "black", size = 5) +
  geom_point(aes(y = presence_mean), fill = "#00A08A", shape = 21, colour = "black", size = 5) + 
#   geom_point(aes(y = total.surv), fill = NA, shape = 21, colour = "black", size = 5,
#              data = subset(pred_present, treatment == "control")) + 
  ylab("Mean survival (n = 5)") + xlab("Predator number") + 
  mytheme

pred_number <- pred_present %>% 
  filter(ispred == "present") %>%
  mutate(npred = ifelse(grepl(" \\+ ", treatment), "two", "one")) 

number <- pred_number %>%
  ggplot(aes(x = npred, y = trtmeans)) +
  geom_point(colour = "black", fill = NA, shape = 21, colour = "black", size = 5) +
  stat_summary(fun.y = mean, fill = "#00A08A", shape = 21, size = 5, geom = "point") +
  ylab("Mean survival (n = 5)") + xlab("Predator presence") + 
  mytheme
  
require(gridExtra)
grid.arrange(pres, number, ncol=2)

```

**Figure 2**:  Adding predators reduces insect emergence, probably because they eat the prey. However, increasing predator number raises survival again. This is because the predators are interfering with each other.

```{r FIG_predator_composition}
pd %>%
  select(treatment, total.surv) %>%
  mutate(treatment = factor(treatment, 
                            levels = c("andro","elong","tabanid","leech",
                                       "elong + andro", "elong + tab", "elong + leech"))) %>%
  filter(treatment != "control") %>% 
  mutate(npred = ifelse(grepl(" \\+ ", treatment),
                        "Two species combination", "Monoculture")) %>%
  ggplot(aes(x = treatment, y = total.surv)) +
  facet_grid(. ~ npred, scales = "free_x") +
  geom_point(colour = "black", fill = NA, shape = 21, colour = "black", size = 5) +
  stat_summary(fun.y = mean, fill = "#00A08A", shape = 21, size = 5, geom = "point") +
  mytheme
```

**Figure 3**: Predator community compositon, in monoculture and polyculture, and its effects on prey survival. Prey emergence is lowest when damselflies are paired together, but increases when damselflies are combined with other species.

```{r FIG_PD_experiment_nonadditive,fig.cap=""}

summarize_randoms_phylo %>% 
  filter(response_var == "survival") %>%
  ggplot(aes(x=phylodistance,y=mean)) + 
  geom_linerange(aes(ymin=lower, ymax=upper)) + 
  geom_point(size=6,shape = 21, colour = "black", fill = "#00A08A") + 
  ylab("Mean treatment difference, Control-Treatment") + 
  xlab("Time (Mya)") + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey") + 
  mytheme
# 
# 
# ggplot(summarize_randoms_phylo,
#        aes(x=Time,y=mean))+geom_errorbar(aes(ymin=lower, ymax=upper),width=0)+geom_point(size=3)+ylab("Mean treatment difference, Control-Treatment")+xlab("Time (Mya)")+facet_wrap(~variable)
```

**Figure 4**: Combinations of predators beyond congenerics show a negative non-additive effect on predation rate.  Points represent the mean difference between the means of two monocultures compared to the mean of a polycultures. The dashed line represents differences $= 0$, i.e. no nonadditive effect. These values are presented as differences from (predator free) controls; therefore zero indicates no significant non-additive effect. Relative to control (no predator) plants, bromeliads containing two predators which were not congeneric showed less predation


```{r summaryFs}
pd2 <- pd %>% 
  select(treatment, total.surv, decomp, fine, N, growth) %>%
  mutate(treatment = treatment %>% 
           equals("control") %>% 
           ifelse(paste0(treatment,
                         seq_along(treatment)),
                  treatment)) %>% 
  gather(key = response, value = resp_val, -treatment) %>%
  mutate(ispred = ifelse(grepl("control[0-9]+", treatment), "absent", "present"),
         npred = ifelse(grepl(" \\+ ", treatment), "two", "one"),  
         # note this mislabels control; remember to filter it
         species_pair = ifelse(treatment == "elong + andro",
                               "Leptagrion.andromache_Leptagrion.elongatum", NA),
         species_pair = ifelse(treatment == "elong + leech",
                               "Leptagrion.elongatum_Tabanidae.spA", species_pair),
         species_pair = ifelse(treatment == "elong + tab",
                               "Leptagrion.elongatum_Tabanidae.spA", species_pair)) %>%
  left_join(phylogenetic_distance)
         


trtnames <- c("total.surv" = "Total prey survival",
  "fine" = "FPOM (g)",
  "decomp" = "Decomposition (g)",
  "growth" = "Bromeliad growth")

mean_na <- function(x) mean(x, na.rm = TRUE)


modlist <- list(
  presence = pd2 %>% 
    select(treatment, response, resp_val, ispred) %>%
    group_by(ispred, treatment, response) %>% 
    summarise(mean = mean_na(resp_val)) %>% 
    group_by(response) %>% 
    do(m = lm(mean ~ ispred, data = .)),
  identity = pd2 %>% 
    filter(npred == "one" & ispred != "absent") %>%
    select(treatment, response, resp_val, ispred) %>% 
    group_by(response) %>%
    do(m = lm(resp_val ~ treatment, data = .)),
  number = pd2 %>% 
    filter(ispred == "present") %>%
    select(treatment, response, resp_val, npred) %>%
    group_by( response, npred, treatment) %>%
    summarise(mean = mean_na(resp_val))  %>% 
    group_by(response) %>%
    do(m = lm(mean ~ npred, data = .)),
  phylodist = pd2 %>%
    filter(npred == "two") %>%
    group_by(response) %>%
    do(m = lm(resp_val ~ phylodistance, data = .))
  )

prF <- function(resp,test,.modlist = modlist){
  modsum <- .modlist[[test]] %>%
    filter(response == resp) %>%
    extract2("m") %>%
    extract2(1) %>%
    summary 
  
  fv <- modsum %>%
    extract2("fstatistic")
  F_format <- paste0("F~",fv[["numdf"]],",",fv[["dendf"]],"~", " = ")
  pv <- pf(fv[1],fv[2],fv[3],lower.tail=F)
  
  fval <- round(fv[["value"]], 2)
  
  if (pv < 0.05) {
    paste0(F_format,"**",fval,"**")
    } else {
      paste0(F_format,fval)      
    }
  }

```

**Table 1** Experimental results. We measured for community-level variables: total prey survival (both emerged adults and surviving larvae), the breakdown of coarse detritus, the production of fine particulate organic matter (FPOM), and the growth of the bromeliad itself. We considered our experimental design at several levels: comparing treatments with predators to those without, contrasting predator species, comparing predator communities of 1 or 2 species, and considering the effects of phylogenetic distance between predators.

Response | Predator Presence | Identity | Richness | Pairwise PD |
-------- | ------------------| ---------|--------- | ---------- |
Total prey survival  |   `r prF("total.surv", "presence")` | `r prF("total.surv", "identity")`  | `r prF("total.surv", "number")` | `r prF("total.surv", "phylodist")`
Decomposition (g)  | `r prF("decomp", "presence")` | `r prF("decomp", "identity")`| `r prF("decomp", "number")` | `r prF("decomp", "phylodist")`
FPOM (g)     | `r prF("fine", "presence")` | `r prF("fine", "identity")` | `r prF("fine", "number")` | `r prF("fine", "phylodist")`
Bromeliad growth  | `r prF("growth", "presence")` |`r prF("growth", "identity")` | `r prF("growth", "number")` | `r prF("growth", "phylodist")`


## Discussion

In our system, distantly related predators are no more likely to be similar in distribution than are close relatives (Question 1). We also tested whether diet similarity decreased with predator phylogenetic distance (Question 2), and found a stronger pattern: diet similarity declined with phylogenetic distance between two predators, in a nearly linear relationship. However, this difference in diet did not translate into a difference in the composition of animals surviving in bromeliads stocked with different predators (3a). We did find an effect of phylogenetic distance when predators were placed in pairs - specifically, when damselflies were paired with non-damselfly predators. This resulted in more prey animals surviving to the end of the experiment; this could be caused by a reduction in feeding rate when predator taxa are very dissimilar. 

### Do related organisms occur in different bromeliads? 

Predator pairs showed only a weak relationship between phylogenetic distance and differences in distribution. Such differences could have been caused by different species either having different preferences for environmental variables, or antagonistic interactions with each other. The absence of any signal suggests that either patches do not differ in variables which matter to the predators, or that predator taxa do not have strong impacts on each other's distribution. The lack of a relationship is probably not due to a lack of variation in bromeliad communities: We know already (CITE) that bromeliads in this and other systems vary considerably in many habitat variables, such as detritus content, amount of sunlight, and habitat size. It is more likely that these predator species are habitat generalists. Organisms which live in small, fluctuation-prone habitats -- especially if they are long-lived -- may evolve a wide physiological tolerance to those environmental fluctuations. Therefore we would not expect a high degree of habitat specificity among these organisms. 

Although we found some evidence for negative intraspecific effects in our experiment, these do not appear to limit predator distribution. This could be caused by a low encounter rate of predators in natural plants, especially in larger bromeliads than those used in our experiment. Additionally, negative effects could still be occurring in nature, but not resulting in mortality. Indeed, if animals are actually able to adjust their development times and feeding rates when conditions are unfavourable (and if the presence of a competing predator creates such unfavourable conditions), then we would expect to observe more frequent, not less frequent, co-occurances of predators.  Non-consumptive negative interactions need not result in dissimilar distributions.

In our experimental treatments with paired predators we observed little predator mortality; this suggests that predator indirect interactions might reduce predator feeding rates but do not necessarily result in predator mortality. Our observational data indicate that at the level of the patch (i.e. a single bromeliad) a wide range of predator phylogenetic diversity is possible, from very similar to very disparate. This justifies the phylogenetic diversities we used in our experimental communities, as these are within the range found in nature.

### Do related organisms have similar diets?

Predators consumed very similar prey, except those most phylogenetically distinct. There were some slight taxon-specific diet preferences which accounted for this. Leeches and tabanids showed different patterns of consumption compared with *Leptagrion* predators: they consumed less frequently, and notably consumed prey in less than half of trials with *Culex* (leech) or *Scirtes* (Tabanid).  This may be because these predators lack the strong jaws and "mask" of odonata, and these two prey species are particularly active (in the case of culicids) or difficult to handle (scirtids).  Tabanids in particular differ in microhabitat use from odonates; living deep within leaf axils, where culicids are rarely found. One of our *Leptagrion* morphospecies showed a greater preference for harder-bodied prey species (i.e. Ostracoda, Scirtes and *Phylloicus*, which is a caddisfly.)  All of these animals are consumed more frequently by this damselfly than by the other species of _Leptagrion_, while the other damselflies more frequently consumed *Culex* and *Polypedilum* prey. Thus, our data show some evidence of a phylogenetic basis for diet dissimilarity, based on the very different traits of these invertebrate taxa. Traits can be more important than phylogeny *per se* to a predator's diet: [@Moody1993] found that unrelated decapod species which were morphologically similar were also functionally similar.
Similarly, [@Rezende2009] found that both body size and phylogeny determined the food web "compartment" (shared predator-prey interactions) of a predator in a marine foodweb. In addition, our experimental results are consistent with high similarity among predator diets: all predator species had comparable effects on all response variables, including prey species survival.

### Phylogenetic diversity and non-additive effects

Predator pairs which were more phylogenetically distant had a larger negative nonadditive effect on prey capture.  This is contrary to our hypothesis that more distant predators would show an increase in prey capture via complementarity. *L. andromache* did not produce a negative effect in combination with *L. elongatum*, while more phylogenetically distinct predators did. It may be that these odonates have behavioural traits that reduce the probability of their interaction, for example each nymph occupying a single leaf-well. If this is the case, each damselfly may not experience many cues indicating the presence of other predators, resulting in no non-additive interaction.

Odonates can be sensitive to chemical cues of potential predators, which causes a decrease in feeding rate (Barry and Roberts 2014). However other research in bromeliads has demonstrated that it is physical contact with other organisms that reduces damselfly predation rate (Trish).  If this is the case for *Leptagrion*, then when combined with leeches they may be responding to frequent contact with those very active predators. Tabanids, however, are rarely observed outside of a deep leaf axil -- in this case, it may be chemical cues which are responsible. There may also be a phylogenetic signal to the chemical cues which the damselflies perceive: i.e. close relatives (other _Leptagrion_ sp. in this case) might induce less of an effect than other predators.

Interestingly, the pattern of induced defenses do not line up with predation risk, as observed in our feeding trials. In feeding trials with leeches, actual predation was rarely observed -- except in one instance, where the damselfly ate the leech. However, trait-mediated indirect effects may not always reflect realized probability of predation: an animal may change behaviour when exposed to a "predator" which poses little threat (REF).

One limitation of our approach is the focus on a single focal predator, the odonate *Leptagrion elongatum*, which was common in all of our predator treatments. It is possible that this species is more sensitive to the presence of other predators, and therefore shows a larger trait-mediated indirect effect, than would other species in this community.  However, this is the most common species in this community and our results indicate that its top-down effects are likely to be frequently reduced by the presence of other predators.

## References