% Predator phylogenetic diversity decreases predation rate via antagonistic interactions
% A. Andrew M. MacDonald; Diane S. Srivastava; Gustavo Q. Romero
%

```{r options_packages_data, message=FALSE, echo=FALSE}
## set chunk options
opts_chunk$set(message=FALSE,warning=FALSE,echo=FALSE,fig.cap=FALSE,dev="png")

## load packages
library(ggplot2)
library(bipartite)
library(reshape2)
library(vegan)
library(picante)
library(beanplot)
library(pander)
library(gplots)
library(plyr)
library(gridExtra)

# occurrance data -- abundance
occur <- read.csv("../data/reorganized_data/predator.cooccur.txt",stringsAsFactor=FALSE)
# occurrance data -- metabolic capacity
metabolic  <- read.csv("../data/reorganized_data/predator.cooccur.metabolic.txt",
                       stringsAsFactor=FALSE)
# feeding trial data
foodweb <- read.csv("../data/reorganized_data/reorganized.feeding.trial.data.csv",
                    stringsAsFactors=FALSE)
# experimental data
pd <- read.csv("../data/reorganized_data/pd_exp_cleaned_data.csv",stringsAsFactors=FALSE)
# enriched leaves -- ie just the N data for the detritus we put in.
enriched <- read.csv("../data/reorganized_data/enriched_leaves.csv",stringsAsFactors=FALSE)
# phylogeny data
predtree_timetree_ages <- read.tree("../data/reorganized_data/predator_tree_time.newick")
# randomization tests
rand.means <- read.csv("../data/predator.div.experiment/randomizations.group.means.csv")

## load in functions
source("../R.scripts/FUNCTIONS_predator_diversity.R")

#source("../R.scripts/cleanup_organization_for_graphing_analysis.R")
```

```{r experiment_data_postprocess}
#dropping a record that seems to have been 90% decomposed!
pd[which(pd$decomp>0.7),'decomp'] <- NA
## just to make sure:
#with(pd,table(treatment))
#head(pd)

```

## Introduction

* A large literature documents the various effects of predators on one another, and the consequences of these interaction for communities; meanwhile studies which measure (and occasionally manipulate) phylogenetic diversity of producers are increasingly common -- yet we know very little about the ecosystem-level effects of predator phylogenetic diversity.

* predators are important for ecosystem processes, and for the diversity and abundance of their prey species; mecahanisms of predator diversity effect include complementarity, intra-guild predation, competition.

* Phylogenetic methods have been useful in understanding the effects of species combinations in experiments with other functional groups; a similar approach might help to synthesize the diverse literature on predator diversity effects.

* Phylogenetic information can provide predictive insight beyond a system, and also can allow us to make comparisons between partly-overlapping or nested sets of species interactions, by supplying a common measure of dissimilarity between a wide range of species

* Observational datasets can provide clues to negative associations o-occurrance; phylogenetic diversity has proven espeically useful in these cases 

* The phylogenetic diversity of a foodweb might reflect niche conservatism among predators -- that is, closely related predators might eat similar diets. 

<!-- Additionally, the structure of a foodweb determines how it will respond to changes in diversity
-->

* The top-down effects of predators, alone and in combinations, can be measured at the scale of an entire ecosystem by using mesocosm experiments.

<!--  Predators are present in most ecosystems, and are important functional
groups in determining ecosystem function.  While predator-prey relationships
have been studied for a long time, we understand little of the effects of
predator diversity on whole communites and ecosystems.  Phylogenetic diversity
of plants may correlates well with community level variables, but as yet studies
of predator combinations rarely use measures of predator phylogenetic diversity.
In our study we present results from patterns of predator co-occurance, feeding
trials, finally a community-lveel experiment in which we directly manipulated
phylogenetic diversity of predators.  In each we ask if the phylogenetic
distance between predators is related to similarity, or if diversity is
correlated with effect on ecosystem function . Decreasing predator richness has
been shown to increase herbivory [@Byrnes2006] in a three-level kelp food web.
As these authors point out, the effect of diversity on ecosystem functioning is
better known for lower tropic levels, rather than predators.    Predator
combinations can have many different outcomes.  From the perspective of
ecosystem function it is important to consider whether these result in more or
less top-down control.  Predator effects can be direct via changes in
consumption, indirect via non-consumptive effects.  in other words, it can be
via the effects of predators on each other, or on their prey, and directly or
indirectly.  Therefore, in our experiment we tracked both predator and prey
survival to the end of the experiment

for example, predators can kill each other, or decrease feeding rates. 


We hypothesize that the phylogenetic relationships among predatory taxa in this
system can be used to interpret their ecology. Specifically, we test the
hypothesis that phylogenetic relatedness is negatively correlated with
probability of co-occurance, positively with diet similarity.  Consequently, we
might predict that ecosystem function peaks at some intermediate level of
phylogenetic diversity -- where predators occur but where their similarity
creates complementarity.

-->

* We test three related hypotheses concerning co-occurance patterns, diet similarity and top-down ecosystem effects of diverse predators, using a natural mesocosm: the community of invertebrates living within bromeliads. 

1. *species co-occurance*: closely-related predators occur together more
frequently than less-related predators, due to their similar habitat
requirements.  Additionally, very closely related species never co-occur because
they are  too similar.

2. *diet similarity*: similarity in diet (as measured by feeding trials)
decreases with phylogenetic distance.

3. *ecosystem-level effects*: similarity in the effect of predators on whole
ecosystems declines with phylogenetic distance.  Additionally, the non-additive
effect of predators will have a greater absolute value when their phylogenetic
diversity is larger.


## Methods

## Site and Species

We conducted all observations and experiments in Parque Estadual da Ilha do
Cardoso ($25^{\circ} 03^{\prime}$ S, $47^{\circ}53^{\prime}$ W), a 22.5 ha
island off the south coast of São Paulo state, Brazil. We worked in a closed
coastal forest (restinga) the understory of which is mostly covered by
_Quesnelia arvensis_ Mez. (Bromeliaceae), a large terrestrial bromeliad that
accumulates up to 2.8 L of rainwater in tanks formed by individual leaves. More
than 47 species of macroinvertebrates are found in these aquatic communities.
This diversity encompasses an entire foodweb, from detrital shredders
(Tipulidae, Trichoptera), detrital scrapers (Scirtidae), collectors
(Chironomidae, Syrphidae, Psychodidae), and filter feeders (Culicidae) to a
diverse predator assemblage consisting of at least 3 species of _Leptagrion_
(Zygoptera:Coenagrionidae), 2 species of predatory fly (Tabanidae), and 2
species of leech (Hirudinidae).  Many other smaller or occasional predators also
occur, such as predatory midges (Chironomidae:Tanypodinae) and Dytiscid beetles.


###  Observations of predator co-occurance 

We used an observational dataset collected in 2008 to examine patterns of co-
occurance among predator taxa.  This dataset was collected by completely
dissecting bromeliads and washing all leaves to remove invertebrates.  The
resulting water was filtered through two sieves (250 and 850 µm), which removed
particulate organic matter without losing any invertebrates.  All invertebrates
were counted and identified to the lowest taxonomic level possible. They were
also measured, when possible, or placed in size categories.

Because predators in this system vary widely in size, we scaled our measure of
coexistence by the predator's overall effect on the community.  We used
"metabolic capacity", which correlates with predator feeding rate [@Brown2004;
@Wilby2005].  We calculated body mass using allometric equations which related
dry biomass to body length.   Metabolic capacity is calculated as biomass raised
to the power of 0.69; this reflects the nonlinearity of feeding rate on body
size across many invertebrate taxa.  We calculated total metabolic capacity of
each predator species in each bromeliad by summing estimates for all individuals
of the same species within a plant.

### Diet similarity

We quantified diet similarity between predators in a series of feeding trials.
We  covered all potential predator-prey pairs present in the experiment
(described below), and attempted to perform all other trials whenever possible.
We were limited by the density of animals we could find in the field, with some
predators being rarer than others.

We conducted our feeding trials between March and April 2011.  We placed
predators together with prey in a 50ml vial, with a leaf or stick for
substrate. The only exception to this setup was the tabanid larvae, which we
placed between two vertical surfaces to imitate the narrow space found in
bromeliad leaf axils.  Generally our trials contained a single predator and a
single prey individual, except in the case of very small prey (_Elpidium_ sp.)
or predators (_Monopelopia_ sp.) in which case we increased the density.  We
replicated each combination up to 5 times where possible, and allowed 1 day for
predation to occur.

### Community effect experiment

Our experiment was created with two goals in mind: first, to measure the major
effects of these predators on their prey and second, to estimate the non-
additive effects of pairwise predator combinations.  The strength of these non
additive effects can then be related back to the phylogenetic distance between
each member of a pair of predators.

We selected the most common predators in this system, which also created a range
of relatedness: two congeneric damselflies (*Leptagrion andromache* and
*Leptagrion elongatum*), two insects (*L. elongatum* and a Tabanid predatory
fly), and two invertebrates (*L. elongatum* and leeches).  We used each of the
pairs of predators just described, as well as each species in monoculture, and a
predator-free control (8 treatments, n=5).  Combinations were substitutive,
maintaining the same amount of predator metabolic capacity in each.  Response
variables included the rate of decomposition of leaves, bromeliad growth and
insect emergence.  This experiment allows the estimation of the effect of each
predator species (monoculture treatments), as well as the detection of non-
additive effects in predator combinations.

In Feburary 2011, bromeliads between 90 and 200ml were collected, thoroughly
washed and soaked for 12 hours in a tub of water.  They were then hung for 48
hours to dry.  One bromeliad dissected after this procedure contained no
insects.

Each bromeliad was supplied with dried leaves, simulating natural detritus
inputs from the canopy.  We enriched these leaves with N-15 by fertilizing five
(Jabuticaba, *Plinia cauliflora*) plants with 40ml/pot/day of 5g/L ammonium
sulphate containing 10 percent atom excess of N15. *duration*. started on
27/1/2011 Whole leaves were then picked from plants and air-dried until constant
weight, and then soaked for three days and the water discarded. About 1.5 g of
leaves were placed in each bromeliad (`r mean(pd$mass.g.)` ± `r sd(pd$mass.g.)`).

Each bromeliad was stocked with a representative insect community. The densities
of each prey taxon were calculated from the 2008 observational dataset, using
data from bromeliads of similar size to those in our experiment.  All densities
used were within the range of these calculated abundances, and all experimental
bromeliads received the same insect community.  Halfway through the experiment,
insects were added to bromeliads a second time to simulate continuous
oviposition.  After addition of the prey community, all bromeliads were enclosed
with a mesh cage topped with a malaise trap and checked daily for emergence of
adults.

In this experiment we measured five response variables: production of fine
particulate organic matter (FPOM), decomposition of coarse detritus, bromeliad
growth, cycling of Nitrogen into plant tissue, and survival of prey (emergence +
surviving larvae). We analyzed each of these responses with ANOVA.  We
calculated the non-additive response as follows: the difference between the
polyculture mean (n=5) and the mean of both monoculture means for each predator
(n=5 for each monoculture).  We generated bootstrap confidence intervals for
these nonadditive effects; confidence intervals which do not overlap zero
indicate a significant nonadditive effect of a predator combination.

<!--
\begin{table}
  \centering
  \caption{densities of each species}
  \label{tab:sppden}
  \begin{tabular}{l l}
    \hline
    \emph{Chironomus detriticula} & 10 \\
    \emph{Polypedium} sp. 1 & 4 \\
    \emph{Polypedium} sp. 2 & 2 \\
    \emph{Psychodid} sp. 1 & 1 \\
    \emph{Scyrtes} sp. A & 5 \\
    \emph{Culex} spp. & 4 \\
    \emph{Trentepholia} sp. & 1
  \end{tabular}
\end{table}
-->

## Results

```{r CALC_distance_matrices}
####### metabolic matrix ####
## we need to calculate two distance matrices:
## 1) metabolic capacity distance
## 2) phylogenetic distance

## metabolic matrix -- the "distance" between predator co-occurance, measured as metabolism
## remove the first column -- it's species names
metabolic.matrix <- metabolic[-1]
## put that name column as dimnames
dimnames(metabolic.matrix)[[1]] <- metabolic[,1]

## now that metabolic capacity is set up, there are several ways for us to go forward: 
## euclidian distance, or maybe correlations?

## euclidian distance between metabolic densities
pred.abd.distance <- vegdist(metabolic.matrix,method="euclid")
occur_matrix <- as.matrix(pred.abd.distance) # convert to matrix

## correlations between metabolic densities
metabolic_mat <- as.matrix(metabolic.matrix)
# reordered metabolic distance matrix
#occur_matrix <- cor(t(metabolic_mat))

####### phylogeny matrix ####
## Calculate distances
allpred_phylodist <- cophenetic(predtree_timetree_ages)

######## 
```

```{r CALC_diet_similarity}
# Check for TRUE ZEROS in cast matrix.

# trial.list <- split(foodweb,foodweb$predator.names)
# sapply(trial.list,nrow)
## need predators as columns, herbivores as rows
foodweb.cast <- dcast(data=foodweb,formula=Prey.species~predator.names,value.var="eaten.numeric",fun.aggregate=sum)
# remove species names
foodweb.matrix <- as.matrix(foodweb.cast[,-1])
# have better names
dimnames(foodweb.matrix) <- list(foodweb.cast[[1]],names(foodweb.cast)[-1])
foodweb.matrix <- foodweb.matrix[,-ncol(foodweb.matrix)]  ## last column was an NA predator.

# make the distance matrix -- with the jaccard index?
## finally, calculate distance
distances <- vegdist(t(foodweb.matrix),method='jaccard',diag=TRUE)
## make a distance matrix so lower.tri subsetting works
diet_dist_mat <- as.matrix(distances)

```

```{r CALC_experiment-randomization_summary}
### ---- group means ####
##go2 <- responses.means(1000)
#write.csv(go2,"randomizations.group.means.csv")
## order these correctly
rand.means$sp.pair <- factor(rand.means$sp.pair,
                             levels=c('elong + andro','elong + tab',
                               'elong + leech')
                             )
## remove the X column
rand.means <- rand.means[,-1]

### supplementary figure?  ####
meansMelt <- melt(rand.means)
# #densityplot(~growth+survival+fine+decomp,groups=sp.pair,data=go)
# ggplot(data=meansMelt,aes(x=value,colour=sp.pair))+geom_histogram()+facet_wrap(~variable)

## summarize the randomizations 
summarize_randoms <- ddply(.data=meansMelt,.variables=.(sp.pair,variable),summarize,
                       mean=mean(value),lower=quantile(value,probs=c(0.025)),
                       upper=quantile(value,probs=c(0.975))
                       )
#####
```

```{r DATA_merge_similarity_matrices}
## we need to merge together several matrices:
## metabolic occurance + predator phylogenetic distance
## diet similarity + predator phylogenetic distance
## experiment randomization results + predator phylogenetic distance

## note that the nomeclature of the columns keeps `sp.pair` as the only shared name among columns.
#### metabolic occurance ####
metabolic_df <- melt(occur_matrix)[melt(upper.tri(occur_matrix))$value,]
names(metabolic_df) <- c("metapred1", "metapred2", "metadistance")
metabolic_df$sp.pair <- paste(metabolic_df$metapred1,metabolic_df$metapred2,sep="_")


## Phylogenetic distance ####
allpred_phylodist_df <- melt(allpred_phylodist)[melt(upper.tri(allpred_phylodist))$value,]
names(allpred_phylodist_df) <- c("phylopred1", "phylopred2", "phylodistance")
allpred_phylodist_df_RH <- allpred_phylodist_df
allpred_phylodist_df_LH <- allpred_phylodist_df
allpred_phylodist_df_RH$sp.pair <- paste(allpred_phylodist_df_RH$phylopred1, allpred_phylodist_df_RH$phylopred2,sep="_")
allpred_phylodist_df_LH$sp.pair <- paste(allpred_phylodist_df_LH$phylopred2, allpred_phylodist_df_LH$phylopred1,sep="_")

## Diet similarity ####
diet_df <- melt(diet_dist_mat)[melt(upper.tri(diet_dist_mat))$value,]
names(diet_df) <- c("dietpred1", "dietpred2", "dietdistance")
diet_df$sp.pair <- paste(diet_df$dietpred1,diet_df$dietpred2,sep="_")
## we also need to know how many trials each predator has in common:
incommon <- t(foodweb.matrix>0)%*%(foodweb.matrix>0)
incommon_df <- melt(incommon)[melt(upper.tri(incommon))$value,]
names(incommon_df) <- c("commonpred1", "commonpred2", "Ncommon")
incommon_df$sp.pair <- paste(incommon_df$commonpred1,incommon_df$commonpred2,sep="_")

#### randomization results ####
# distances of L. elongatum to everything:
Le_distances <- sort(allpred_phylodist["Leptagrion.elongatum",])
## a lookup table to pair spp leves with time-since-divergence
lkup <- data.frame(sp.pair=levels(summarize_randoms$sp.pair),
                   Time=Le_distances[c("Leptagrion.andromache","Tabanidae.spA","Hirudinidae")])

#### merging ####
## metabolic occurance + predator phylogenetic distance
metabolic_occur_phylo1 <- merge(metabolic_df,allpred_phylodist_df_LH)
metabolic_occur_phylo2 <- merge(metabolic_df,allpred_phylodist_df_RH)
metabolic_occur_phylo <- rbind(metabolic_occur_phylo1,metabolic_occur_phylo2)
## diet similarity + predator phylogenetic distance
diet_df1 <- merge(diet_df,incommon_df)

diet_similarity_phylo1 <- merge(diet_df1,allpred_phylodist_df_LH)
diet_similarity_phylo2 <- merge(diet_df1,allpred_phylodist_df_RH)
diet_similarity_phylo <- rbind(diet_similarity_phylo1,diet_similarity_phylo2)
## experiment randomization results + predator phylogenetic distance
#diet_df1$sp.pair[!diet_df1$sp.pair%in%diet_similarity_phylo$sp.pair]
## as this code shows, small lepts are not in the running yet
summarize_randoms_phylo <- merge(summarize_randoms,lkup)

#####
```

### metabolic capacity and phylogenetic distance

```{r TEST_metabolic~phylo,results='asis'}
meta_phylo_lm <- with(metabolic_occur_phylo,lm(metadistance~phylodistance))
meta_phylo_lm_summary <- summary(meta_phylo_lm)
```

```{r nodecalc}
nodeages <- lapply(list.files(path="../data/TreeData/",pattern="*.csv",full.names=TRUE),read.csv)
names(nodeages) <- list.files(path="../data/TreeData/",pattern="*.csv")
nstudies <- sapply(nodeages,nrow)
#names(nodeages)[which(nstudies>1)]
n.nodes <- length(nodeages)
```

Within the 2008 observational dataset, we identified `r length(predtree_timetree_ages$tip.label)` species as predators. These predators
vary in taxonomic relatedness: from congeners -- _Bezzia_ sp.
(Diptera:Ceratopogonidae) with two species and _Leptagrion_ sp.
(Odonata:Coenagrionidae) with three --  to confamilials (three species of
Tabanidae and two of Empididae, all Diptera). Three families of Diptera are
represented by a single species each: Dolichopodidae, Corethrellidae and
Chironomidae. The deepest taxonomic divide is between all insects present and a
species of leech (Annelida:Hirudinidae). Node age data was available for all but
the shallowest nodes of the tree, where either a lack of taxonomic information
(e.g. Tabanidae) or a lack of phylogenetic study (e.g. _Leptagrion_) prevented
more information from being included.  These branches were left as polytomies,
and were all assigned identical, arbitrary and short branch lengths (15 Mya).

We obtained node age estimates for all `r n.nodes` internal nodes of the tree.  These were usually provided by only a single study, with more studies available for deeper nodes: Insecta--Hirudina (`r paste(range(nodeages[["insects.to.leeches.csv"]]$Time),collapse=" to ")` Mya, n=`r nstudies[["insects.to.leeches.csv"]]` studies), Odonata--Tabanidae (`r paste(range(nodeages[["odonata-Tabanidae.csv"]]$Time),collapse=" to ")` Mya, n=`r nstudies[["odonata-Tabanidae.csv"]]` studies) and Tabanidae--Diptera (`r paste(range(nodeages[["odonata-Tabanidae.csv"]]$Time),collapse=" to ")` Mya, n=`r nstudies[["tabanidae_culidicae_ie_Diptera.csv"]]` studies).  We used the median estimate of age for these nodes.

In 2008, insects were counted and measured in an observational study of `r ncol(metabolic.matrix)` bromeliads. Across all bromeliads, predator species differed widely in metabolic capacity, from `r min(rowSums(metabolic.matrix))` for a species of Empididae, to `r max(rowSums(metabolic.matrix))` for the abundant predator _Leptagrion andromache_.  Predators often co-occured in bromeliads ($`r mean(colSums(metabolic.matrix>0))` \pm `r round(sd(colSums(metabolic.matrix>0)),digits=2)`$ species per plant).  However, the euclidian distance between the total metabolic capacity of two predators did not show any relationship with phylogenetic distance between them (F~`r meta_phylo_lm_summary$fstatistic[["numdf"]]`,`r meta_phylo_lm_summary$fstatistic[["dendf"]]`~=`r meta_phylo_lm_summary$fstatistic[["value"]]`, p=`r round(pf(meta_phylo_lm_summary$fstatistic[1],meta_phylo_lm_summary$fstatistic[2],meta_phylo_lm_summary$fstatistic[3],lower.tail=FALSE),digits=2)`).

### diet similarity and phylogenetic distance
```{r dietcalc}
foodweb_len <- dcast(data=foodweb,formula=Prey.species~predator.names,value.var="eaten.numeric",fun.aggregate=length)
# remove species names
foodweb_len_mat <- as.matrix(foodweb_len[,-1])
# have better names
dimnames(foodweb_len_mat) <- list(foodweb_len[[1]],names(foodweb_len)[-1])
foodweb_len_mat <- foodweb_len_mat[,-ncol(foodweb_len_mat)]  ## last column was an NA predator.

## what percentage of total trials resulted in predation?
percentpredation <- round(colMeans(foodweb.matrix/foodweb_len_mat,na.rm=TRUE),digits=2)

```

We conducted `r sum(foodweb_len_mat)` feeding trials of `r ncol(foodweb.matrix)` predator taxa fed `r nrow(foodweb.matrix)` prey taxa.  However, due to the rarity of some taxa many predator-prey pairs were not possible; we tested `r sum(foodweb_len_mat>0)` pairwise combinations.  Most trials were replicated at least 5 times, but the number of replicates for various combinations ranged from `r min(foodweb_len_mat[foodweb_len_mat>0])` to `r max(foodweb_len_mat[foodweb_len_mat>0])`.  Two damselflies, _Leptagrion andromache_ and _Leptagrion elongatum_, showed the higest rates of prey consumption (prey consumed in `r percentpredation[["Leptagrion.andromache"]]*100`% and `r percentpredation[["Leptagrion.elongatum"]]*100`% of trials, respectively). 

```{r TEST_diet-sim_as_PD}
diet_phylo_lm <- with(diet_similarity_phylo,lm(dietdistance~phylodistance,weights=Ncommon))
diet_phylo_lm_summary <- summary(diet_phylo_lm)
## test a squared term with 
```

All predators showed a very generalist diet breadth, consuming nearly all species offered to them.  However, more phylogenetically distant predators preferred slightly different diets, as measured by euclidian distance between feeding trial outcomes (F~`r diet_phylo_lm_summary $fstatistic[["numdf"]]`,
`r diet_phylo_lm_summary $fstatistic[["dendf"]]`~=`r round(diet_phylo_lm_summary $fstatistic[["value"]],2)`, p=`r round(pf(diet_phylo_lm_summary $fstatistic[1],diet_phylo_lm_summary $fstatistic[2],diet_phylo_lm_summary $fstatistic[3],lower.tail=FALSE),digits=3)`)  Regression was weighted by the number of trials conducted.

### Ecosystem-level effects and phylogenetic distance

```{r experimentsummary}
predeffect <- function(resp="total.surv"){
  diffeffect <- (mean(pd[[resp]][pd$treatment!="control"],na.rm=TRUE)-mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE))/mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE)
  round(diffeffect,digits=2)*100
  }
# 
# ddply(pd,.(treatment),summarize,meansurv=mean(total.surv))
# 
# mean(pd$emerged)
# mean(rowSums(pd[c("Culicidae","Chironomidae","Tipulidae","Scirtidae")])
```

In our manipulative experiment, we placed a standardized prey community into bromeliads and measured five response variables: the total survivorship (both emerged during experiment and found as larvae at the end) of all prey, the %N15 which was transferred into bromeliad tissue, bromeliad growth, coarse detritus decomposition and fine organic matter production.  Predators had a large effect on prey survivorship: on average all predator treatments showed `r abs(predeffect())`% lower prey emerging or surviving as larvae relative to the predator-free control.  Nitrogen transport to bromeliad leaves was slightly decreased relative to controls (`r predeffect("N")`%), and was only higher than the control in treatments including Tabanid predators.  We found a similar pattern for plant growth: on average, predators had a `r predeffect("growth")`% effect on growth of bromeliad leaves (mm), though Tabanids seemed to create a slight increase. The decomposition of coarse detritus and production of fine organic matter showed no obvious pattern related to the mere presence of predators.

```{r polysummary}
polyeffect <- function(resp="total.surv"){
  diffeffect <- (mean(pd[[resp]][pd$treatment%in%c("elong + andro","elong + leech","elong + tab")],na.rm=TRUE)-mean(pd[[resp]][pd$treatment%in%c("andro","tabanid","leech","elong")],na.rm=TRUE))/mean(pd[[resp]][pd$treatment=="control"],na.rm=TRUE)
  round(diffeffect,digits=2)*100
  }

```

Predator combinations tended to have an increased effect on our response variables relative to predators alone.  Approximately `r polyeffect()`% more prey survived in polyculture, on average,  compared to all monocultures.  Effects were smaller for Nitrogen (`r polyeffect("N")`%) and bromeliad growth (`r polyeffect("growth")`%). Fine particulate organic matter was produced `r polyeffect("fine")`% more when predators were present in combination.

Our experimental design allows us to estimate the non-additive effect of predator species pairs on whole communities of prey, and the functioning of the bromeliad ecosystem.  We used randomization tests to test the hypothesis that increased phylogenetic distance between members of a predator pair results in a greater magnitude of nonadditive effect.  We contrasted the differences of the mean individual predator treatments from the control with the mean difference of their pairwise combination from the control.  We found the greatest effect for prey survival: while effects of _L. andromache_ and _L. elongatum_ in combination were quite similar to the effect of either alone, when _L. elongatum_ was placed in the same plant as either a Tabanid larva or leeches, on average 5 more prey (18% of total prey community) survived till the end of the experiment (Fig 3).  This effect was smaller among the other variables, most of which showed confidence intervals from the randomization test which overlap 0.


### Figures

```{r FIG_metabolic_occurance_as_phylo,fig.cap="Figure 1: Phylogenetic distance and predator co-occurance.  Each point represents a pair of predator species.  We caluclated total metabolic capacity for each predator species in each bromeliad, and then calculated co-occurance between two predators as the euclidian distance between total metabolic capacity of two species."}
ggplot(metabolic_occur_phylo,aes(x=phylodistance,y=metadistance))+geom_point()+xlab("phylogenetic distance")+ylab("euclidian distance between total metabolic capacity")
```


```{r FIG_feeding_trial_as_phylo, fig.cap="Figure 2: Phylogenetic distance and diet similarity.  We performed 237 feeding trials with the 8 major predator taxa found in this system.  We found that more distantly-related predators consume more dissimiliar prey.  We measured diet similarity as euclidian distance among feeding trial outcomes; this measure includes information about the number of predation events in each predator-prey pairing.  Regression was weighted by the sample size of the predator-prey pair."}
ggplot(data=diet_similarity_phylo,aes(y=dietdistance,x=phylodistance,size=Ncommon))+geom_point()+scale_size(range=c(2,6))+ylab("Euclidian distance between feeding trials")+xlab("phylogenetic distance")+stat_smooth(method="lm", aes(weight=Ncommon))
# plot(dist.mat[lower.tri(dist.mat)]~
#        jitter(phylodist[lower.tri(phylodist)],amount=10),
#      xlab="phylogenetic distance",ylab="jaccard distance between feeding trials")
```


```{r FIG_experiment_responses,fig.cap="Figure 3: Treatment responses for the manipulative experiment.  Points are means, and horizontal lines represent minimum and maximum values for each response.  n=5 for all means."}
pd_long <- melt(pd[names(pd)%in%c("treatment","total.surv","fine","decomp","growth","N")],id.vars="treatment")

plotmaker <- function(resp,kill_trtnames=TRUE,label){
ggplot(pd_long,aes(y=value,x=treatment))+stat_summary(fun.y = mean, fun.ymin = min, fun.ymax = max,geom="pointrange",subset=.(variable==resp))+geom_hline(x=0,colour="grey")+geom_hline(x=mean(subset(pd_long$value,pd_long$treatment=="control"&pd_long$variable==resp),na.rm=TRUE))+ylab(label)+coord_flip()+if(kill_trtnames) theme(axis.text.y = element_blank(), axis.title.y=element_blank())
}

surv <- plotmaker(resp="total.surv",kill_trtnames=FALSE,label="prey survival")
N <- plotmaker("N",label="Nitrogen")
growth <- plotmaker("growth",label="growth (mm)")
decomp <- plotmaker("decomp",label="decomposition \n (g)")
fine <- plotmaker("fine",label="production of \n FPOM (g)")
grid.arrange(surv,N,growth,decomp,fine,ncol=5,widths=unit(c(2,rep(1, 4)), "null"))

##  ADD COLOUR FOR THE TYPE: CONTROL (LARGER, BLACK), SINGLE AND POLY.  NO LEGEND. 
```


```{r FIG_PD_experiment_nonadditive,fig.cap="Figure 4: Combinations of predators beyond congenerics show a negative non-additive effect on predation rate.  Points represent the mean difference between the means of two monocultures compared to the mean of a polycultures.  These values are presented as differences from (predator free) controls; therefore 0 indicates no significant non-additive effect. Relative to control (no predator) plants, bromeliads containing two predators which were not congeneric showed less predation"}
ggplot(subset(summarize_randoms_phylo,summarize_randoms_phylo$variable=="survival"),
       aes(x=Time,y=mean))+geom_errorbar(aes(ymin=lower, ymax=upper),width=0)+geom_point(size=3)+ylab("Mean treatment difference, Control-Treatment")+xlab("Time (Mya)")
# 
# ggplot(summarize_randoms_phylo,
#        aes(x=Time,y=mean))+geom_errorbar(aes(ymin=lower, ymax=upper),width=0)+geom_point(size=3)+ylab("Mean treatment difference, Control-Treatment")+xlab("Time (Mya)")+facet_wrap(~variable)
```


## Discussion

### co-occurrance
* Predators appear to be generalist with regard to their habitat preference; indicating that prey face a heterogeneous landscape of predator risks, and demonstrating that our experimental design captures natural variation in predator composition.

* Predators occuring together at smaller spatial scales (i.e. within the same plant) may have limited direct interactions due to habitat structure, for example via the physical obstructions caused by detritus

### diet similarity
* We demonstrate a slight but important decrease in feeding preferences with phylogenetic distance, indicating that food web structure itself might vary with phylogenetic diversity, becoming more compartmentalized when predators are more dissimilar.

* The differences in predation rates and diet breadth among our predators represents important trait differences between the groups involved, including differences in metabolic rate and feeding mode. (_for example, Tabanids and leeches are peircing predators, while Leptagrion is gape-limited._)

<!--In fact, a better analysis for taht data might be to compare the relative probability of eatin a certain prey species for leeches, and compare to same for Leptagrion.  In other words: if Leptagrion, leeches, and tabanids had eaten the same number as they had originally, but these were randomly distributed among all prey -- would we see the same pattern which we found?  or a stronger or weaker one?

weighting feeding-trial regression by the metabolic capacity of predators, rather than sample size?  Does it correlate with feeding rate?  marginal distribution / rank-abd plot showing variation in metabolic capacity among predators?

A mechanistic model for the effects on the whole community, elegantly tied together with the results of the previous sections -- the differences in feeding rate, diet breadth, but no difference in habitat preference.  
-->

* Feeding differences between predators, combined with their patchy but unpredictable distribution among bromeliads, creates a large variation in food web structure and total amount of predation per patch within this bromeliad metacommunity 

### community experiment

* Phylogenetically diverse predator assemblages showed more prey survivorship (i.e. less predation) than expected; this may be to strong trait-mediated indirect effects on the feeding rate of _Leptagrion elongatum_ when in the presence of other non-damselfly predators.

* In response to predator manipulations we observed less transfer of Nitrogen into bromeliad tissue than did Ngai and Srivastava (2006); this may be due to a greater role for bacterial decomposition in this system (cite?).

## References