---
title: "Appendix"
author: "Andrew MacDonald"
output:
  pdf_document: default
  md_document: default
  html_document: default
  word_document:
    reference_docx: reference.docx
bibliography: /home/andrew/Documents/Mendeley_reference_lists/@MS_pdpaper.bib
---

```{r results='asis', echo=FALSE,message=FALSE}
library(magrittr)
library(dplyr)
library(pander)
library(picante)
```

### predator phylogeny

```{r}
read.tree("../data/reorganized_data/predator_tree_time.newick") %>% plot
```



### Prey community composition
Densities of prey species used in the experiment.

|  Species                 | density |
|:------------------------:|:-------:|
| *Chironomus detriticula* |  10     |
| *Polypedium sp. 1*       |  4      |
| *Polypedium sp. 2*       |  2      |
| *Psychodid sp. 1*        |  1      |
| *Scyrtes sp. A*          |  5      |
| *Culex spp.*             |  4      |
| *Trentepholia sp.*       |  1      |

### Model comparison for distributional similarity as a function of PD
```{r}
formulae <- c("quadratic" = "$a \\times (PD)^2 + b \\times PD + c$",
              "bellshaped" = "$peak \\times {e}^{(-1 \\times (PD)^2 / a)}$",
              "{e}^{onential" = "$b \\times {e}^{(a \\times PD)}$",
              "Sshaped" = "$\\frac{c \\times {e}^{(a \\times PD)}}{(c \\times {e}^{(a \\times PD)} + (1 - c))}$",
              "linear" = "$a \\times x + b$",
              "constant" = "a \\times x$"
              )
```


```{r results='asis',echo=FALSE}

read.csv(file = "distributional_similarity.csv") %>%
  mutate(Equation = formulae[model]) %>%
  select(Equation,AIC) %>%
  pander
```


### Model comparison for diet similarity as a function of PD
Note: these models are weighted by the number of prey species tested
```{r results='asis',echo=FALSE}


read.csv(file = "diet_similarity.csv") %>%
  mutate(Equation = formulae[model]) %>%
  select(Equation,AIC) %>%
  pander
```

