---
title: "Supplementary Material"
author: "A. Andrew M. MacDonald; Diane S. Srivastava; Gustavo Q. Romero"
output:
  pdf_document: default
  md_document: default
  html_document: default
  word_document:
    reference_docx: reference.docx
bibliography: /home/andrew/Documents/Mendeley_reference_lists/@MS_pdpaper.bib
---

```{r results='asis', echo=FALSE,message=FALSE}
library(magrittr)
library(dplyr)
library(pander)
library(picante)
library(stargazer)
library(tidyr)
library(knitr)
opts_chunk$set(message=FALSE,warning=FALSE,echo=FALSE,fig.cap=FALSE)

```

### predator phylogeny

```{r}
read.tree("../data/reorganized_data/predator_tree_time.newick") %>% plot
```


### Model comparison for distributional similarity as a function of PD
```{r results='asis',echo=FALSE}
formulae <- c("quadratic" = "$a \\times (PD)^2 + b \\times PD + c$",
              "bellshaped" = "$peak \\times {e}^{(-1 \\times (PD)^2 / a)}$",
              "{e}^{onential" = "$b \\times {e}^{(a \\times PD)}$",
              "Sshaped" = "$\\frac{c \\times {e}^{(a \\times PD)}}{(c \\times {e}^{(a \\times PD)} + (1 - c))}$",
              "linear" = "$a \\times x + b$",
              "constant" = "a \\times x$"
              )

read.csv(file = "distributional_similarity.csv") %>%
  mutate(Equation = formulae[model]) %>%
  select(Equation,AIC) %>%
  pander
```

### Model comparison for diet similarity as a function of PD
Note: these models are weighted by the number of prey species tested
```{r results='asis',echo=FALSE}


read.csv(file = "diet_similarity.csv") %>%
  mutate(Equation = formulae[model]) %>%
  select(Equation,AIC) %>%
  pander
```

## Predator diversity experiment

### Prey community composition
Densities of prey species used in the experiment.

|  Species                 | density |
|:------------------------:|:-------:|
| *Chironomus detriticula* |  10     |
| *Polypedium sp. 1*       |  4      |
| *Polypedium sp. 2*       |  2      |
| *Psychodid sp. 1*        |  1      |
| *Scyrtes sp. A*          |  5      |
| *Culex spp.*             |  4      |
| *Trentepholia sp.*       |  1      |

### experimental summary
```{r exp summary, results='asis'}
read.csv("../data/reorganized_data/pd_exp_cleaned_data.csv",
         stringsAsFactors=FALSE) %>%
  tbl_df() %>%
  select(treatment,total.surv,fine,decomp,growth,N) %>%
  gather(response,value,-treatment) %>%
  group_by(treatment,response) %>%
  summarise(meanval=mean(value),
            n=n(),
            sdev=sd(value)) %>%
  mutate(SE=sdev/sqrt(n),
         meanval_sig=signif(meanval,2),
         SE_sig=signif(SE,2),
         meanSE=paste(meanval_sig,"±",SE_sig)) %>%
  select(treatment,response,meanSE) %>% 
  spread(response,meanSE) %>%
  stargazer(summary = FALSE, type = "latex")

# # could rename with this
# coltext <- expression("surviving insects","fine detritus (g)","leaf decomposition (g)",
#                       "bromeliad growth (g)","Nitrogen cycling")
```

### the effect of predator diversity *per se*
```{r experimentANOVA,results='asis'}
pd <- read.csv("../data/reorganized_data/pd_exp_cleaned_data.csv",
               stringsAsFactors=FALSE) %>%
  tbl_df() %>%
  transform(decomp=ifelse(decomp>0.7,NA,decomp))

### calculate means 
trt_means <- pd %>%
  select(treatment,total.surv,fine,decomp,growth,N) %>%   
  gather(response,value,-treatment) %>%
  group_by(treatment,response) %>%
  summarise(meanval=mean(value,na.rm = TRUE),
            n=n()) %>%
  mutate(trt_type = treatment %>% grepl(".+ \\+ .+",.) %>% ifelse(.,"pair",treatment),
         trt_type = trt_type %in% c("control","pair") %>% not %>% ifelse(.,"single",trt_type)) %>%
  ungroup %>%
  group_by(trt_type,response) %>%
  arrange(response,trt_type) %>%
  filter(trt_type != "control") %>%
  select(trt_type,response,val = meanval)

## select control treatments
cntrl_val <- pd %>% 
  select(treatment,total.surv,fine,decomp,growth,N) %>%   
  gather(response,value,-treatment) %>%
  filter(treatment == "control") %>%
  select(trt_type = treatment, response, val = value)

## order treatments & do model
models <- rbind(trt_means,cntrl_val) %>%
  arrange(response) %>%
  group_by(response) %>%
  mutate(repno = seq_along(trt_type)) %>%
#   spread(response,val) %>%
  mutate(trt_type = trt_type %>% ordered(x = .,levels = c("control","single","pair"))) %>%
  arrange(response,trt_type) %>%
  select(-repno) %>%
  do(mod = lm(val ~ trt_type, data = .))

## present via stargazer
models %>%
  extract2("mod") %>%
  set_names(models$response) %>%
  stargazer(.,type = "latex",column.labels = models$response %>% as.character)

```
